### 2021-04-22 Notes Modelleren-B 20-04-2021.tex
Initial Overleaf Import

\documentclass{article}
\usepackage[utf8]{inputenc}

\title{Notes Modelleren-B 20-04-2021}
\author{Severijn Wiechers}
\date{April 2021}

\begin{document}

\maketitle

\section*{member list}

\begin{itemize}
    \item Severijn Wiechers 4948262
    \item Bart Stolk 5414946
    \item Dyami C.M. Silvério 5172497
    \item Wessel Van Sommeren 5007712
\end{itemize}

\section*{meeting schedule}

Aan het eind van iedere meeting wordt een schema opgesteld van de onderdelen die we als groep af willen hebben voor het begin van de volgende meeting. De meetings zijn in principe op de ingeroosterde uren.

\section*{Rough planning}
lijst van alle deadlines:
\begin{itemize}
    \item 23-04-2021: Eerste opzet van afspraken
    \item 12-05-2021: peer review
    \item 03-06-2021: draft report
    \item 10-06-2021: feedback draft report
    \item 15-06-2021: Pitch
    \item 17-06-2021: Research report
    \item 18-06-2021: Evaluation report 
\end{itemize}
Algemene planning:

In de eerste weken(4.2, 4.3, 4,4, 4.6) ligt de focus op het ontwikkelen van de python file en het algoritme. Week 4.7 voornamelijk ingedeeld voor het afronden van de code en het schrijven van het draft report. In de week na het draft report wordt er tijdens de meeting gekeken naar het feedback geven op de andere verslagen. In week 4.9 wordt de feedback verwerkt in het draft report om tot een eindverlag te komen. Verder zullen hier ook de taken van de ptich worden verdeeld.

\section*{five criteria for peer review}
    \begin{itemize}
       \item Punctuality
       \item Motivation / Participation
        \item Adaptability
        \item Quality of the work 
        \item Communication
    \end{itemize}

\section*{Evaluation report}
\begin{itemize}
    \item Introduction
    \item Planning
    \item Project management
    \item File management and communication
    \item Quality
\end{itemize}

\section*{Quality plan}
In het quality plan worden alle algemene afspraken geschreven voor het opleveren en opslaan van het werk.

\begin{itemize}
    \item Het verslag wordt geschreven in latex/overleaf
    \item De vergaderingen zullen plaatvinden in de groepen van teams
    \item Python files worden aan gewerkt via git
    \item Alle bestanden worden opgedeeld en opgeslagen per onderwerp (denk hierbij aan notulen, verslag of bijvoorbeeld .py files)
    \item De naamgeving van bestanden wordt als volgt gegeven: "Hoofd onderwerp - datum - titel van bestand - maker"

\end{itemize}
\end{document}

### 2021-04-28 Minutes Modelleren-B Week 4.2.tex

--- 2021-04-22 Notes Modelleren-B 20-04-2021.tex
@@
 
-\title{Notes Modelleren-B 20-04-2021}
-\author{Severijn Wiechers}
+\title{Notule Week 4.2}
+\author{Modelleren-B groep 15}
 \date{April 2021}
@@
 \begin{document}
+Functies maken:
+-Functie van Excel naar Python omzetten
+-Cost function
 
-\maketitle
-
-\section*{member list}
-
-\begin{itemize}
-    \item Severijn Wiechers 4948262
-    \item Bart Stolk 5414946
-    \item Dyami C.M. Silvério 5172497
-    \item Wessel Van Sommeren 5007712
-\end{itemize}
-
-\section*{meeting schedule}
-
-Aan het eind van iedere meeting wordt een schema opgesteld van de onderdelen die we als groep af willen hebben voor het begin van de volgende meeting.
-De meetings zijn in principe op de ingeroosterde uren.
-
-\section*{Rough planning}
-lijst van alle deadlines:
-\begin{itemize}
-    \item 23-04-2021: Eerste opzet van afspraken
-    \item 12-05-2021: peer review
-    \item 03-06-2021: draft report
-    \item 10-06-2021: feedback draft report
-    \item 15-06-2021: Pitch
-    \item 17-06-2021: Research report
-    \item 18-06-2021: Evaluation report 
-\end{itemize}
-Algemene planning:
-
-In de eerste weken(4.2, 4.3, 4,4, 4.6) ligt de focus op het ontwikkelen van de python file en het algoritme.
-Week 4.7 voornamelijk ingedeeld voor het afronden van de code en het schrijven van het draft report.
-In de week na het draft report wordt er tijdens de meeting gekeken naar het feedback geven op de andere verslagen.
-In week 4.9 wordt de feedback verwerkt in het draft report om tot een eindverlag te komen.
-Verder zullen hier ook de taken van de ptich worden verdeeld.
-
-\section*{five criteria for peer review}
-    \begin{itemize}
-       \item Punctuality
-       \item Motivation / Participation
-        \item Adaptability
-        \item Quality of the work 
-        \item Communication
-    \end{itemize}
-
-\section*{Evaluation report}
-\begin{itemize}
-    \item Introduction
-    \item Planning
-    \item Project management
-    \item File management and communication
-    \item Quality
-\end{itemize}
-
-\section*{Quality plan}
-In het quality plan worden alle algemene afspraken geschreven voor het opleveren en opslaan van het werk.
-
-\begin{itemize}
-    \item Het verslag wordt geschreven in latex/overleaf
-    \item De vergaderingen zullen plaatvinden in de groepen van teams
-    \item Python files worden aan gewerkt via git
-    \item Alle bestanden worden opgedeeld en opgeslagen per onderwerp (denk hierbij aan notulen, verslag of bijvoorbeeld .py files)
-    \item De naamgeving van bestanden wordt als volgt gegeven: "Hoofd onderwerp - datum - titel van bestand - maker"
-
-\end{itemize}
+Ideeën:
+-
 \end{document}

### 2021-04-28 cost_functie.py
Excel in python

@@ -1,0 +1,11 @@
+import pandas as pd 
+
+xls = pd.ExcelFile('large.xlsx')
+df1 = pd.read_excel(xls, 'w')
+df2 = pd.read_excel(xls, 'c')
+df3 = pd.read_excel(xls, 'f')
+
+
+print(df1)
+print(df2)
+print(df3)

### 2021-04-29 cost_functie.py
Cost

@@ -1,11 +1,47 @@
 import pandas as pd 
 
-xls = pd.ExcelFile('large.xlsx')
-df1 = pd.read_excel(xls, 'w')
-df2 = pd.read_excel(xls, 'c')
-df3 = pd.read_excel(xls, 'f')
+
+file_name ='Large.xlsx'
+class parcel():
+	def __init__(self,file):
+		xls = pd.ExcelFile(file)
+		self.flow = pd.read_excel(xls, 'w')
+		self.package_cost = pd.read_excel(xls, 'c')
+		self.hub_cost = pd.read_excel(xls, 'f')
+		self.cities = len(self.flow.columns)
+
+	def calculate_cost(self,hubs):
+		cost = 0
+		index = {}
+		flow_copy = self.flow
+		cost_copy = self.package_cost
+		cost_copy = cost_copy.take([hub - 1 for hub in hubs])
+		print(cost_copy)
+		for i in range(1,self.cities):
+			if i in hubs:
+				index[i] = i
+				continue
+			nearest_hub = cost_copy.loc[cost_copy[i] > 0 , i].idxmin()
+			cost += cost_copy[i][nearest_hub] * flow_copy[i].sum()
+			nearest_hub = nearest_hub + 1
+			print(i,nearest_hub)
+			index[i] = nearest_hub
+			flow_copy[nearest_hub] = flow_copy[i]+flow_copy[nearest_hub]
+			flow_copy = flow_copy.drop(columns = i)
+		print(index)
+		# for i in range(self.cities):
+		# 	to_hub = index[i]
+		# 	flow_copy.iloc[i]
+
+		print(flow_copy.iloc[14])
 
 
-print(df1)
-print(df2)
-print(df3)
+
+
+test = parcel(file_name) 
+test.calculate_cost([1,2,3,9])
+
+
+print(test.package_cost.loc[test.package_cost[14] > 0, 14].idxmin())
+print(test.package_cost[14])
+print(test.package_cost)

### 2021-04-29 Minutes Modelleren-B Week 4.2.tex
Updates from Overleaf

@@
 \title{Notule Week 4.2}
-\author{Modelleren-B groep 15}
+\author{Modelleren\itemB groep 15}
 \date{April 2021}
@@
 \begin{document}
-Functies maken:
--Functie van Excel naar Python omzetten
--Cost function
+\section*{Notules Woensdag}
+Aanwezig: Iedereen \newline
+Voorzitter: Severijn \newline
+Secretaris: Bart \newline
+\subsection*{Functies maken:}
+\begin{itemize}
+    \item Functie van Excel naar Python omzetten
+    \item Cost function (Wessel)
+\end{itemize}
 
-Ideeën:
--
+\subsection*{Ideeën "intuitief" algoritme:}
+\begin{itemize}
+    \item Hubs die veel flow naar zichzelf hebben
+    \item Grootste kostbepalers onderzoeken
+    \item Flow en kosten combineren in edges
+    \item Waar gaan meeste parcels naartoe?
+    \item Probleem verdelen in kleinere problemen en oplossen.
+    \item Een hub per keer kiezen
+\end{itemize}
+
+\subsection*{Deelproblemen:}
+\begin{itemize}
+    \item Hubs baseren op flow naar zichzelf (Dyami)
+    \item Collectie miminaliseren (Severijn)
+    \item Dsitributie minimaliseren (Bart)
+    \item Hoeveel hubs? (Wessel)
+\end{itemize}
+
+De resultaten van deze deelproblemen combineren om 1 voor 1 hubs af te wegen en kiezen.
+
+\section*{Notules Donderdag}
+\subsection*{feedback week 1}
+Aanwezig: Bart, Dyami, Severijn \newline
+Voorzitter: Severijn \newline
+Secretaris: Bart \newline
+\begin{itemize}
+    \item Preciesere notules (wie is aanwezig, voorzitter, notulist, wie doet wat de volgende keer)
+    \item Schedule voor voorzitter en notulist
+    \item Planning uitbreiden
+    \item Planning Python implementaties
+    \item Eerder beginnen met draft report
+    \item Hoe wordt controle gedaan? Wie checkt wie?
+\end{itemize}
+
+\subsection*{Voorlopige Afspraken}
+Aanwezig: Iedereen \newline
+Voorzitter: Severijn \newline
+Secretaris: Bart \newline
+\begin{itemize}
+    \item Cost functie afmaken - (Wessel) - Voor 4/5/21
+    \item ILP Begrijpen - (Iedereen) - Voor 4/5/21    \item Intuïtief Algoritme - (Wessel, Dyami) - Week 4.3
+    \item ILP Implementatie -  Week 4.3
+    \item ILP Python Implementation - (Severijn, Bart) - Week 4.4
+\end{itemize}
 \end{document}

### 2021-05-25 ILP.tex
Add files via upload
Eerste ILP formulatie

\documentclass[11pt]{article} % use larger type; default would be 10pt
\usepackage{amsmath}
\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)

\usepackage{geometry} % to change the page dimensions
\geometry{a4paper} % or letterpaper (US) or a5paper or....

\title{ILP of Parcel Delivery}
\author{Bart Stolk}

\begin{document}
\maketitle

\section{Input}
\begin{align*}
c_{ij} &= \text{cost between city i and j}  \quad i,j=1,\ldots,N \\
w_{ij} &= \text{flow between city i and j} \quad i,j=1,\ldots,N \\
f_{i} &= \text{cost to turn city i into a hub} \quad i=1,\ldots,N \\
\chi &= \text{collection cost multiplier} \\
\alpha &= \text{transfer cost multiplier} \\
\delta &= \text{distribution cost multiplier} \\
\end{align*}

\section{Variables}
\begin{align*}
h_{i} &=
\begin{cases} 
1, & \text{if city i is a hub} \quad i=1,2,\ldots,N \\ 
0, & \text{otherwise} 
\end{cases} \\ 
e_{ij} &=
\begin{cases} 
1, & \text{if there is and edge between city i and j} \quad i,j=1,2,\ldots,N \\ 
0, & \text{otherwise} 
\end{cases} \\
n_{ij} &= 
\begin{cases} 
0, & \text{if } h_{i} = 1 \text{ and } h_{j}=1 \quad i,j=1,2,\ldots,N \\ 
e_{ij}, & \text{otherwise}
\end{cases}
\end{align*}

\section{Objective function}
$$min\,z=\sum^N_{i=1}\sum^N_{j=1}\sum^N_{k=1}\sum^N_{l=1}(w_{ij}(\chi c_{ik}e_{ik}+(\alpha c_{kl}+\delta c_{lj})e_{lj})) + \sum^N_{i=1}f_{i}h_{i}$$
s.t.
\begin{align*}
e_{ij}-h_{i}-h_{j} &= 0 \quad i,j=1,2,\ldots,N \text{ for } i \ne j \\
e_{ii}-h_{i} &= 0 \quad i=1,2,\ldots,N \\
\sum^N_{j=i}n_{ij} &= 1 \quad i=1,2,\ldots,N \\
\sum^N_{i=1}h_{i} &\geq 1 \quad i=1,2,\ldots,N
\end{align*}
\end{document}

### 2021-05-25 ILP.tex
Update ILP.tex
Updated the ILP to use correct variables (and some formatting changes)

@@
 \begin{align*}
-c_{ij} &= \text{cost between city i and j}  \quad i,j=1,\ldots,N \\
-w_{ij} &= \text{flow between city i and j} \quad i,j=1,\ldots,N \\
-f_{i} &= \text{cost to turn city i into a hub} \quad i=1,\ldots,N \\
+c_{ij} &= \text{cost between city i and j}  &i=1,2,\ldots,N \quad& j=1,2,\ldots,N\\
+w_{ij} &= \text{flow between city i and j} &i=1,2,\ldots,N \quad& j=1,2,\ldots,N\\
+f_{i} &= \text{cost to turn city i into a hub} &i=1,2,\ldots,N \quad\\
 \chi &= \text{collection cost multiplier} \\
@@
 \begin{cases} 
-1, & \text{if there is and edge between city i and j} \quad i,j=1,2,\ldots,N \\ 
+1, & \text{if there is and edge between city i and j} \quad i=1,2,\ldots,N \quad j=1,2,\ldots,N  \\ 
 0, & \text{otherwise} 
 \end{cases} \\
-n_{ij} &= 
+y_{ij} &=
 \begin{cases}
-0, & \text{if } h_{i} = 1 \text{ and } h_{j}=1 \quad i,j=1,2,\ldots,N \\ 
-e_{ij}, & \text{otherwise}
-\end{cases}
+1, & \text{if i is a non-hub city and has an edge with city j} \quad i=1,2,\ldots,N \quad j=1,2,\ldots,N \\
+0, & \text{otherwise} 
+\end{cases} \\
+x_{ij} &=
+\begin{cases}
+1, & \text{if i is a non-hub city and has an edge with a hub j} \quad i=1,2,\ldots,N \quad j=1,2,\ldots,N \\
+0, & \text{otherwise} 
+\end{cases} \\
+z &= \text{total cost of the given hubs and connections}
 \end{align*}
@@
 \begin{align*}
-e_{ij}-h_{i}-h_{j} &= 0 \quad i,j=1,2,\ldots,N \text{ for } i \ne j \\
-e_{ii}-h_{i} &= 0 \quad i=1,2,\ldots,N \\
-\sum^N_{j=i}n_{ij} &= 1 \quad i=1,2,\ldots,N \\
-\sum^N_{i=1}h_{i} &\geq 1 \quad i=1,2,\ldots,N
+e_{ii}-h_{i} &= 0 &i=1,2,\ldots,N \quad&\\
+\sum^N_{j=1}y_{ij} &= 1 &i=1,2,\ldots,N \quad&\\
+\sum^N_{j=1}x_{ij} &=1 &i=1,2,\ldots,N \quad&\\
+y_{ij} &\leq (1-h_i) &i=1,2,\ldots,N \quad& j=1,2,\ldots,N \\
+y_{ij} &\leq e_{ij} &i=1,2,\ldots,N \quad& j=1,2,\ldots,N \\
+y_{ij} &\geq e_{ij}-h_i &i=1,2,\ldots,N \quad& j=1,2,\ldots,N \\
+x_{ij} &\leq y_{ij} &i=1,2,\ldots,N \quad& j=1,2,\ldots,N \\
+x_{ij} &\leq h_j &i=1,2,\ldots,N \quad& j=1,2,\ldots,N \\
+x_{ij} &\geq y_{ij} + h_j - 1 &i=1,2,\ldots,N \quad& j=1,2,\ldots,N \\
+\sum^N_{i=1}h_{i} &\geq 1 &i=1,2,\ldots,N \quad& \\
+h_i, e_{ij}, y_{ij}, x_{ij} &\in \{0,1\} &i=1,2,\ldots,N \quad& j=1,2,\ldots,N \\
 \end{align*}

### 2021-05-27 ILP.py
ILP.py
Niet een werkende versie, uitkomst is -1

@@ -1,0 +1,54 @@
+from pulp import *
+import pandas as pd
+import cplex 
+col = 3
+trns = 1
+dist = 2 
+N = range(15)
+file_name ='Large.xlsx'
+xls = pd.ExcelFile(file_name)
+flow = pd.read_excel(xls, 'w')
+p_cost = pd.read_excel(xls, 'c')
+hub_cost =[13530] + list(pd.read_excel(xls, 'f')[13530])
+flow = flow.values[:,1:]
+p_cost = p_cost.values[:,1:]
+
+model = LpProblem('ParcelTransport', sense=LpMinimize)
+
+isHub = LpVariable.dicts('isHub', N, 0, 1, LpBinary)
+
+
+hasEdge = LpVariable.dicts("hasEdge", [(i,j) for i in N
+                                       for j in N],
+                           0, 1, LpBinary)
+y = LpVariable.dicts("y", [(i,j) for i in N
+                                       for j in N],
+                           0, 1, LpBinary)
+
+x = LpVariable.dicts("x", [(i,j) for i in N
+                                       for j in N],
+                           0, 1, LpBinary)
+
+model += lpSum(flow[i][j]*(col*p_cost[i][k]*hasEdge[i,k]+
+                           (trns*p_cost[k][l]+dist*p_cost[l][j])*hasEdge[l,j])
+               for i in N
+               for j in N
+               for k in N
+               for l in N)
++ lpSum(hub_cost[i]*isHub[i] for i in N)
+
+for i in N:
+    model += hasEdge[(i,i)] - isHub[i] == 0
+    model += lpSum(y[(i,j)] for j in N) == 1
+    model += lpSum(x[(i,j)] for j in N) == 1
+    for j in N:
+        model += y[(i,j)] <= 1-isHub[i]
+        model += y[(i,j)] <= hasEdge[(i,j)]
+        model += y[(i,j)] >= hasEdge[(i,j)] - isHub[i]
+        model += x[(i,j)] <= y[(i,j)]
+        model += x[(i,j)] <= isHub[j]
+        model += x[(i,j)] >= y[(i,j)] + isHub[j] - 1
+model += lpSum(isHub[i] for i in N) >= 1
+
+sol = model.solve(CPLEX_CMD())
+print(LpStatus[sol])

### 2021-05-27 draft_report.tex
draft

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{authblk}

\title{Draft Report Modelleren-B}
\author{Bart Stolk \and Dyami Silvério \and Wessel van Sommeren \and Severijn Wiechers}
\date{03 June 2021}

\begin{document}

\maketitle

\newpage
\section*{Summary}
\addcontentsline{toc}{section}{Summary}

\newpage
\tableofcontents

\newpage
\section{Introduction}
Reducing costs is a valuable tool for any package delivery cooperation to stay relevant and make profits. This report will try to show how it is possible to create an algorithm that gives an optimum 
\newpage
\section{Intuitive Algorithm}

    \subsection{Calculating the cost for multiple hubs}
    \subsection{The different algorithms used}
    \subsection{Result of the combined algorithms}

\newpage  
\section{The ILP solver}
We are given a set N of nodes where we can locate a hub. The cost of opening a hub at node i $\in$ N is $F_i$ and the cost of transporting a parcel to node j from node i is $c_{i_j}$, i $\in$ N, j $\in$ N. We need to decide on which nodes we shall open a hub, and from which hub the nodes will be connected. The goal is to minimize the total cost. 
    \subsection{ILP implementation}
    \subsection{IlP result}

\newpage
\section{Expanded intuitive algorithm}

\newpage
\section{Conclusion}

\section{Appendix} 
\end{document}

### 2021-05-27 draft_report.tex
Folders

@@
 \section{Introduction}
-Reducing costs is a valuable tool for any package delivery cooperation to stay relevant and make profits. This report will try to show how it is possible to create an algorithm that gives an optimum 
+
+% Reducing costs is a valuable tool for any package delivery cooperation to stay relevant and make profits.
+This report will try to show how it is possible to create an algorithm that gives an optimum case of hubs such that the total cost is minimized % we moeten de hoofdvraag nog even correct opstellen.
+
+Reducing costs is a valuable tool for any package delivery cooperation to stay relevant and make profits.
+Hereby, algorithms are a useful and necessary tool to optimize a given data set.
+In this case the goal of the report is to find an algorithm to minimize the cost of a package delivery company.
+Hereby a data set is used where the different transfer, collection and distribution costs are given per city.
+The objective of the algorithm is to find the cities where hubs should be build such that the total cost of sending and receiving packages can be brought down to a minimum.
+An important restraint in this matter is that packages can only move between hubs and cities and in particular not directly between cities themselves.
+
+Since this type of optimization problem is a common practice for companies, there already exists an ILP solver which can give the exact hubs for which have the lowest total cost, However, it has shown that these solvers can take up and lot op time and processing power.
+Therefore it is preferable to get an idea for which routes and hubs will give the best, or at least a proper solution to the problem.
+This will be done by creating different intuitive algorithms which will, if done correctly, give similar results using different methods.
+
+In the first chapter the intuitive algorithms will be explained along with the function which is used to calculate the total costs of choosing a specific combination of hubs.
+Thereafter, the chapter will set forth the input and result of the ILP Solver.
+The section in chapter 4 will focus on a reflection between the result of the intuitive algorithm and the solution of the ILP solver.
+In the following chapter multiple expansions of the problem and intuitive algorithm will be given such that a better result is obtained.
+Finally, the last chapter will contain a conclusion of the report.
+
 \newpage
 \section{Intuitive Algorithm}
+As mentioned in the introduction, it can be useful to use an intuitive algorithm in stead of a direct solver for the problem.
+With a self made script it is possible to get a feel for the cities in which building a hub can have a major impact on reducing the cost of the project.
+Before a code can be written that tries to optimize the building of hubs in cities, it is necessary to have a solid view of the costs that a specific combination of hubs will bring forth.
+Therefore a cost-function is written that will give a total cost as output with a combination of hubs as input.
 
     \subsection{Calculating the cost for multiple hubs}
+        To calculate the cost of using specific hubs we first choose a way to determine connections from hubs to cities.
+The most intuitive way to determine those connections was to link each city with the hub that was cheapest to collect and distribute to.
+Intuitively this made sense as it probably the closest.
+If we take for example the Netherlands this would mean that a packet send from Groningen would be picked up by a hub close to Groningen and not by a hub in for example Maastricht.
+So the first step in the cost function algorithm was to go trough the list of cities and determine which hub from the list of inputted hubs was cheapest to collect from for each city.
+When this connection is determined the sum of the list multiplied by the collection factor is added to the cost and the the amount of packages are added to the packages of the hub city.
+This connection between the city and collection hub was stored in memory as it would be used later to determine transfers and delivery.
+As soon as this first step is complete we end up with a Dataframe with as columns the hub cities and rows the packages that need to go to each city.
+The next step is to go row by row and look up what connection that city had with which hub and transfer all the packages of that row to the correct hub again multiplying the amount times the correct cost associated with transfering from hub to hub.
+Then the total of that row at the correct hub will be delivered to the city and thus multiplying the total number of packages by the cost to deliver from that hub to the city and the delivery factor.
+Once that is done for all rows, all packages will be delivered.
+Now we only need to add the cost of building the hubs and we have the total cost of using specific hubs.
+
     \subsection{The different algorithms used}
+
+    After obtaining a function which calculates the cost of assigning hubs to cities, an intuitive algorithm can be made which tries to give an approximation of the combination of hubs that will result in the lowest possible amount of money that the company will have to spend when moving packages.
+    Since the intuitive algorithms are inherently less accurate than a standardized solver, the di
+
+\subsubsection{Hubs self-flow}
+Transport to the same node is free if the node is a hub.
+This algorithm will be looking at the node with the most flow of parcels to itself and make a hub of this node.
+For this we will use the Large data assignment.
+\paragraph{The Algorithm:}
+Node 14 has the most flow of parcels to itself (181), so this becomes our first hub.
+There is only one hub so all other nodes will be connected to this hub.
+The Total cost will be calculated using our cost-function in python.
+The total cost is 182725,0.
+
+The second node with the most flow of parcels to itself is node 2 with 39 parcels.
+Each node will be connected to the hub with the least cost to transport it there.
+For example for node 11 it cost 11 dollars to transport it to hub 14 and 27 dollars to transport it to hub 2, so node 11 will be connected with hub 14.The total cost is 163586
+
+The third node with the most flow of parcels to itself is node 15 with 26 parcels.
+What will happen if this node becomes a hub?
+As before the nodes will be connected to the hub with the least cost to transport it there.
+The total cost is 169342
+
+If there are two nodes with the same amount of parcels to itself the algorithm will choose the node with the most incoming parcels.
+This algorithm will go through all nodes until the total cost by opening an extra hub becomes more expensive then the cheapest situation
+
+\paragraph{Conclusion}
+The cheapest situation here is where there are 5 hubs in nodes 2,6,13,14 and 15.
+We connect each node with the hub where the cost to transport a parcel is the lowest.
+The total cost in this situation would be 161521.
+One can easily see that node 14 must become a hub as it has the most flow of parcels to itself (181).
+In the following subsections we will be looking at the results of the other algorithms and combine those to see if there are similarities.
+
     \subsection{Result of the combined algorithms}

### 2021-05-27 Reports/draft_report.tex
More Folders

--- 2021-05-27 draft_report.tex
@@
     After obtaining a function which calculates the cost of assigning hubs to cities, an intuitive algorithm can be made which tries to give an approximation of the combination of hubs that will result in the lowest possible amount of money that the company will have to spend when moving packages.
-    Since the intuitive algorithms are inherently less accurate than a standardized solver, the di
+    Since the intuitive algorithms are inherently less accurate than a standardized solver, the different aspects of moving packages are split up and solved using different types of algorithms.
+The total process is divided into the collection from the cities to the hubs, fo
 

### 2021-05-31 ILP/ILP.py
Update ILP.py
Almost there, only the objective function does not seem to function correctly. Added some comments for clarity.

@@ -1,54 +1,77 @@
-from pulp import *
 import pandas as pd
-import cplex 
+import pulp as p
+import numpy as np
+#Input
 col = 3
 trns = 1
-dist = 2 
-N = range(15)
-file_name ='Large.xlsx'
-xls = pd.ExcelFile(file_name)
+dist = 2
+cities = range(15)
+
+file = 'Large.xlsx'
+xls = pd.ExcelFile(file)
+
 flow = pd.read_excel(xls, 'w')
-p_cost = pd.read_excel(xls, 'c')
-hub_cost =[13530] + list(pd.read_excel(xls, 'f')[13530])
 flow = flow.values[:,1:]
-p_cost = p_cost.values[:,1:]
 
-model = LpProblem('ParcelTransport', sense=LpMinimize)
+cost = pd.read_excel(xls, 'c')
+cost = cost.values[:,1:]
 
-isHub = LpVariable.dicts('isHub', N, 0, 1, LpBinary)
+hub_cost = pd.read_excel(xls, 'f')
+hub_cost = hub_cost.values[:,1]
+hub_cost = np.append([13530], hub_cost)
 
+#Creating the model
+Parcel = p.LpProblem('ParcelDelivery', p.LpMinimize)
 
-hasEdge = LpVariable.dicts("hasEdge", [(i,j) for i in N
-                                       for j in N],
-                           0, 1, LpBinary)
-y = LpVariable.dicts("y", [(i,j) for i in N
-                                       for j in N],
-                           0, 1, LpBinary)
+#Variables
+h = p.LpVariable.dicts('isHub', cities, cat='Binary')
 
-x = LpVariable.dicts("x", [(i,j) for i in N
-                                       for j in N],
-                           0, 1, LpBinary)
+e = p.LpVariable.dicts('hasEdge', ((i,j) for i in cities
+                                         for j in cities), cat='Binary')
 
-model += lpSum(flow[i][j]*(col*p_cost[i][k]*hasEdge[i,k]+
-                           (trns*p_cost[k][l]+dist*p_cost[l][j])*hasEdge[l,j])
-               for i in N
-               for j in N
-               for k in N
-               for l in N)
-+ lpSum(hub_cost[i]*isHub[i] for i in N)
+x = p.LpVariable.dicts('Helper1', ((i,j) for i in cities
+                                         for j in cities), cat='Binary')
 
-for i in N:
[…]

### 2021-06-01 ILP/ILP.py
Update ILP.py

@@ -32,10 +32,17 @@ e = p.LpVariable.dicts('hasEdge', ((i,j) for i in cities
 x = p.LpVariable.dicts('Helper1', ((i,j) for i in cities
                                          for j in cities), cat='Binary')
 
+y = p.LpVariable.dicts('Helper2', ((i,j,k,l) for i in cities
+                                          for j in cities
+                                          for k in cities
+                                          for l in cities), cat='Binary')
+
+a = p.LpVariable.dicts('Helper4', ((i,j) for i in cities
+                                         for j in cities), cat='Binary')
 #Objective function
 
-Parcel += p.lpSum(flow[i][j]*(col*cost[i][k]*e[(i,k)]+
-                           (trns*cost[k][l]+dist*cost[l][j])*e[(l,j)])
+Parcel += p.lpSum(flow[i][j]*(col*cost[i][k]+
+                           trns*cost[k][l]+dist*cost[l][j])
                for i in cities
                for j in cities
                for k in cities
@@ -53,21 +60,28 @@ for i in cities:
     Parcel += p.lpSum(e[(i,j)] for j in cities) >= 1
     #Every city is collected to at least one other city.
     
-    Parcel += p.lpSum(x[(i,j)] for j in cities) == 1
-    #Every city is connected to exactly one hub. (Connections between hubs are
-    #implied in the objective function.)
+    Parcel += p.lpSum(x[(i,j)] for j in cities) <= 1
+    #Every city is connected to exactly one hub.
+    
     for j in cities:
         Parcel += x[(i,j)] <= e[(i,j)]
-        Parcel += x[(i,j)] <= h[j]
-        Parcel += x[(i,j)] >= e[(i,j)] + h[j] - 1
-        #Helper variable, see lecture notes 2.5
+        Parcel += x[(i,j)] <= (1-h[i])
+        Parcel += x[(i,j)] >= e[(i,j)] + (1-h[i]) - 1
+        #Helper function, see lecture notes 2.5
         
-        Parcel += e[(i,j)] <= (h[i] + h[j])
-        #This mostly makes sure non-hub cities are not connected to eachother.
+        Parcel += a[(i,j)] <= h[i]
+        Parcel += a[(i,j)] <= h[j]
+        Parcel += a[(i,j)] >= h[i] + h[j] - 1
         
+        Parcel += e[(i,j)] >= a[(i,j)]
+        #There must be connections between hubs.
+  
[…]

### 2021-06-01 Intuitive Algoritm/cost_functie.py
Wessel

@@ -1,8 +1,10 @@
 import pandas as pd 
 import matplotlib.pyplot as plt
-collection_factor = 3
+import math
+collection_factor = 1
 transfer_factor = 1
-distribution_factor = 2 
+distribution_factor = 2
+bus_capactity = 50
 
 file_name ='Large.xlsx'
 class parcel():
@@ -29,7 +31,7 @@ class parcel():
 				cost+= self.hub_cost[i]
 				continue
 			nearest_hub = cost_copy.loc[cost_copy[i] > 0 , i].idxmin()
-			cost += cost_copy[i][nearest_hub] * flow_copy[i].sum()*collection_factor
+			cost += cost_copy[i][nearest_hub] * flow_copy[i].sum()*math.ceil(flow_copy[i].sum()/bus_capactity)*collection_factor
 			nearest_hub = nearest_hub 
 			index[i] = nearest_hub
 			flow_copy[nearest_hub-1] = flow_copy[i]+flow_copy[nearest_hub-1]
@@ -41,7 +43,7 @@ class parcel():
 			prices = list(cost_copy[to_hub])
 			total_packages = sum(flowss)
 			transfer_cost = sum([int(flowss[i])*int(prices[i]) for i in range(len(flowss))])
-			distribution_cost = self.package_cost[to_hub][i+1]*total_packages*distribution_factor
+			distribution_cost = self.package_cost[to_hub][i+1]*math.ceil(total_packages/bus_capactity)*distribution_factor*total_packages
 			# print(distribution_cost)13530
 
 			# print(transfer_cost)

### 2021-06-01 Minutes/Minutes Modelleren-B week 4.6 .tex

\documentclass{article}
\usepackage[utf8]{inputenc}

\title{Minutes Modelleren 25-05-2021}
\author{Severijn Wiechers}
\date{May 2021}

\begin{document}

\maketitle
25-05-2021
\section{Attendance}
\begin{itemize}
    \item Wessel was still in the train so he couldn't be in the meeting
    \item Dyami had a doctors appointment but was in the meeting for the majority of the first hour
    \item Bart and Severijn were present the entire meeting
\end{itemize}

\section{what do we need for the draft report}
\begin{itemize}
    \item The complete intuitive algorithm. Here we will write down all the made algorithms and the conclusion of the chosen hubs.
    \item The answer to the problem via the ilp solver

\end{itemize}

\section{What have we done today}
\begin{itemize}
    \item By combining the results of all the intuitive algorithms we think that the combination of the hubs [6, 14, 8] will have one of the lowest possible costs at 144532 euros in total
    \item We still need to get an answer out of the ILP solver, Bart will try to do that today
\end{itemize}

\section{What do we still have to do}
\begin{itemize}
    \item we still have to let the cost function show the different cities that are connected per hub
\end{itemize}

\newpage
27-05-2021
\section{Attendance}
\begin{itemize}
    \item Everyone was were present the entire meeting
\end{itemize}

\section{Draft Report}
The entire layout of the draft report is made. Furthermore, the first two chapters are almost finished. Before the next meeting the goal is to finish the chapters about the intuitive algorithm. Next Tuesday the final chapters of the report will be written.

\section{What have we done today}
\begin{itemize}
    \item We have written the first chapters of the draft report
    \item We have discussed the peer review; there weren't any explicit thing that stood out // that we wanted to discuss further

\end{itemize}

\section{What do we still have to do}
\begin{itemize}
    \item Bart will finish the ILP solver before the next meeting
    \item Every member will finish their part in the report explaining their intuitive algorithm.
\end{itemize}
\end{document}

Hoe verhoud de ilp tot intuitief
crititische blik op de uitkomsten van het algoritme

### 2021-06-01 Reports/draft_report.tex

@@
 \usepackage{authblk}
+\usepackage{graphicx}
+\usepackage[table]{xcolor}
 
@@
     Since the intuitive algorithms are inherently less accurate than a standardized solver, the different aspects of moving packages are split up and solved using different types of algorithms.
-The total process is divided into the collection from the cities to the hubs, fo
+Hereby, the goal is to minimize the error that the functions have by eventually combining multiple intuitive algorithms such that a minimum cost can be achieved.
+The total process is divided into the collection from the cities to the hubs, followed by the transfer between hubs themselves, and as last the distribution of packages from the hubs to the desired cities.
 
-\subsubsection{Hubs self-flow}
+\subsection{Collection of packages}
+
+In this process the cost function will only give an output of the cost that is formed by combining the cost for collection packages.
+To optimize the costs a python script is made such that, when a city is given to be a hub, the code calculates the cost for this specific hub and the cost if the city next in line would be a hub as well.
+If the total cost drops by adding city as a hub, this city will be added to list of cities that are hubs.
+If the cost rises, the city will be discarded from the process.
+If this process is repeated with each time a different starting, a list will appear of possible combination of hubs and there costs.
+To illustrate this process the outputs will be stated below if it were to start with city number 3 as a hub.
+
+\begin{table}[h!]
+\begin{center}
+ \begin{tabular}{||c |c||} 
+ \hline
+ Cost of the combination(€ x 10^3) & combination on cities as hubs \\ [0.5ex] 
+ \hline\hline
+ 169.3 & [3,1]  \\ 
+ \hline
+ 169.8 & [3,2]  \\
+ \hline
+ 169.8 & [3,4]  \\
+ \hline
+ 127.2 & [3,5]  \\ 
+ \hline
+ 132.0 & [3,5,6]  \\
+  \hline
+ 127.4 & [3,5,7]  \\
+ \hline
+ 128.5 & [3,5,8]  \\
+ \hline
+ 124.8 & [3,5,9]  \\
+  \hline
+ 118.7 & [3,5,9,10]  \\
+ [1ex] 
+ \hline
+\end{tabular}
+\end{center}
+\caption{Exampled of the algorithm which adds cities one by one as hubs based on the cost}
+\label{Collection Algorithm}
+\end{table}
+
+Using this method, the algorithm can give the most frequently used cities as hubs.
+To illustrate the difference in accuracy between the different models the table below will give the amount of times cities are used as hubs in each different part.% Herlezen
+
+\begin{table}
+\begin{tabular}{|p{3cm}|p{3cm}|p{3cm}|}
+\hline
+\multicolumn{3}{|c|}{Amount of times cities are used as hubs} \\
+\hline
+City used as hub & Using only collection cost & Using total cost\\
+\hline
+1 &  8 & 5 \\
+\hline
+2 & 4   & 11 \\
+\hline
+3& 7 & 9\\
+\hline
+4    & 1 & 1 \\
+\hline
+5 & 8 & 9\\
+\hline
+6 & 1 & 8   \\
+\hline
+7 & 2& 3 \\
+
+\hline
+8 & 2 & 2 \\
+\hline
+9 &  7& 10 \\
+\hline
+10 & 11 & 11 \\
+\hline
+11 & 1& 2 \\
+\hline
+12 & 1 & 1 \\
+\hline
+13 & 2& 2 \\
+\hline
+14 & 14& 14 \\
+\hline
+\end{tabular}
+\caption{Table of the amount of times each city is used as a hub in every combination}
+\label{amount of times hubs are used in collection} 
+\end{table}
+
+Using this table a conclusion can be drawn which hubs could be used in the optimum configuration.
+However, a big part of cost reduction is the combination of specific hubs.
+Consider city number 6 for example, here it shows that making this city a hub is not a valuable if the algorithm only takes the collection cost in account, but rises significantly if all cost are calculated.
+
+When the algorithm is used for different input combinations based on ~\ref{amount of times hubs are used in collection} the outputs and results show different possible answers with supposedly relative low cost.
+Hereby the algorithm will add a city as hub if this results in a lower cost by the principle given in the beginning of the section.
+
+\begin{table}
+\begin{tabular}{ |p{3cm}|p{3cm}|p{4cm}|}
+\hline
+\multicolumn{3}{|c|}{Cost and combination of hubs with certain input} \\
+\hline
+Input into the algorithm & Output -  collection Cost (€ x 10^3) & Output - total Cost (€ x 10^3)\\
+\hline
+[14] &  92.6 | [14, 1, 3] & 166.4 | [14, 1, 2, 3, 5, 7]  \\
+\hline
+[10] & 90.0 | [10, 2, 14]   & 153.0 | [10, 2, 14, 8] \\
+\hline
+[6, 10]& 93.9 | [6, 10, 14] & 150.4 |  [6, 10, 14]\\
+\hline
+[6, 14] &  82.9 | [6, 14]  &  144.5 | [6, 14, 8] \\
+\hline
+[9, 14] &  100.7| [9, 14, 1, 3] & 160.4 |  [9, 14, 2, 3, 7] \\
+\hline
+[2, 14] & 87.6 | [2, 14, 6] & 156.9 |  [2, 14, 3, 6]\\
+\hline
+[2, 10, 14] & 90.0 | [2, 10, 14] & 153.0 | [2, 10, 14, 8]  \\
+\hline
+\end{tabular}
+\caption{Table of the amount of times each city is used as a hub in every combination}
+\label{amount of times hubs are used in collection} 
+\end{table}
+
+The combination of cities 6, 8 and 14 will probably have one of the lower possible costs for this problem for as well the collection cost as the total cost.
+Using this combination it shows that the cost can be reduced to around 145 thousand euros.
+
+\subsection{Transfer between hubs}
 Transport to the same node is free if the node is a hub.
-This algorithm will be looking at the node with the most flow of parcels to itself and make a hub of this node.
-For this we will use the Large data assignment.
-\paragraph{The Algorithm:}
-Node 14 has the most flow of parcels to itself (181), so this becomes our first hub.
-There is only one hub so all other nodes will be connected to this hub.
-The Total cost will be calculated using our cost-function in python.
-The total cost is 182725,0.
+To optimize the total cost an algorithm is made that will be looking at the node with the most flow of parcels to itself and make a hub of this node.
+After calculating the cost, the next hub with the most flow of parcels to itself will be become a hub.
+If the solution becomes cheaper by adding a hub, the node is added to a list of hubs.
+If the total cost increases, nothing happens and the next node (most parcels to itself) becomes a hub.
+If there are two nodes with the same amount of parcels to itself the algorithm will choose the node with the most incoming parcels.
+Each node will be connected to the hub with the least cost to transport it there.
+For example if it cost 11 dollars to transport from node 11 to node 14 and 27 dollars to transport it to node 2, node 11 will be connected with hub 14.
+This process is repeated until the algorithm goes through all nodes.
+To illustrate we will use the Large data assignment.
 
-The second node with the most flow of parcels to itself is node 2 with 39 parcels.
-Each node will be connected to the hub with the least cost to transport it there.
-For example for node 11 it cost 11 dollars to transport it to hub 14 and 27 dollars to transport it to hub 2, so node 11 will be connected with hub 14.The total cost is 163586
+\begin{table}[h]
+\begin{tabular}{|p{3cm}|p{3cm}|p{3cm}|}
+\hline
+nodes that are hubs & flow of parcels to itself (last node) & Total cost\\
+\hline
+[14] &  181 & 182725 \\
+\hline
+[14,2]   & 39 & 163586 \\
+\hline
+[14,2,15] & 26 & 169342 \\
+\hline
+[14,2,15,6]  & 18 & 156535 \\
+\hline
+[14,2,15,6,13] & 18 & 161521\\
+\hline
+[14,2,15,6,13,4] & 11 & 167037   \\
+\hline
+[14,2,15,6,13,4,3] & 9 & 177126   \\
+\hline
+[14,2,15,6,13,4,3,9] & 9 & 187710    \\
+\end{tabular}
+\end{table}
 
-The third node with the most flow of parcels to itself is node 15 with 26 parcels.
-What will happen if this node becomes a hub?
-As before the nodes will be connected to the hub with the least cost to transport it there.
-The total cost is 169342
-
-If there are two nodes with the same amount of parcels to itself the algorithm will choose the node with the most incoming parcels.
-This algorithm will go through all nodes until the total cost by opening an extra hub becomes more expensive then the cheapest situation
-
-\paragraph{Conclusion}
-The cheapest situation here is where there are 5 hubs in nodes 2,6,13,14 and 15.
-We connect each node with the hub where the cost to transport a parcel is the lowest.
-The total cost in this situation would be 161521.
-One can easily see that node 14 must become a hub as it has the most flow of parcels to itself (181).
+The cheapest situation here is where there are 4 hubs in nodes 2,6,14 and 15.
+We connect each node with the hub where the cost to transport a parcel is the lowest, the total cost in this situation would be 156 thousand dollars.
+One can easily see that it is very profitable to make a hub out of node 14, as it is the node with the most incoming parcels (610).
 In the following subsections we will be looking at the results of the other algorithms and combine those to see if there are similarities.
 
+\newpage
+
     \subsection{Result of the combined algorithms}
+After combining and analysing the results of the different algorithms, a hub will be opened at nodes 14 and 6 as these nodes were also hubs in the cheapest situation of all different algorithms that were used.
+By opening a hub at node 14 the cost for collecting , transferring and distributing can be reduced, as it is the node with the most amount of incoming parcels.
+The total cost in this situation where we only have two hubs is 144870.
+Our intuitive algorithm will then go through the nodes one by one and check which combination (with node 14 and 6) is the cheapest, this node will then be added to the list of hubs.
+This process is repeated until we go through all nodes.
+
+\begin{table}[h]
+\begin{tabular}{||c |c||}
+\hline
+Nodes that are hubs & Total cost\\
+\hline
+[6,14] &  144870 \\
+\hline
+[6,8,14]   & 144532\\
+\hline
+[2,6,8,14] & 148189 \\
+\hline
+[2,5,6,8,14]  & 152593  \\
+\hline
+[2,5,6,8,12,14] & 160386 \\
+\end{tabular}
+\end{table}
+
+The most advantageous situation that was found with the intuitive algorithm is 
+where we make 3 hubs out of nodes 6,8 and 14.
+By doing this the total cost is reduced to 144532 dollars which is 338 dollars cheaper than the previous one.
+(heb dit snel gemaakt verander het als je wilt)
 
@@
 \newpage
-\section{Expanded intuitive algorithm}
+\section{Expanded algorithm}
+The package delivery problem is set out into a data set where only the cost per package and the amount of packages send and received are given.
+To make this problem more realistic a situation is made where all the packages are collected and distributed using a bus.
+In this particular case, a bus can hold up to and including a total of one hundred packages.
+If more packages are needed to move every package, the delivery company will have to send two or more busses.
+
+This scenario is translated into the algorithm by letting all factors of moving packages
 

### 2021-06-03 ILP/ILP.py
Update ILP.py
IT WORKS!

@@ -24,66 +24,45 @@ hub_cost = np.append([13530], hub_cost)
 Parcel = p.LpProblem('ParcelDelivery', p.LpMinimize)
 
 #Variables
-h = p.LpVariable.dicts('isHub', cities, cat='Binary')
+e = p.LpVariable.dicts('hasEdge', ((i,k) for i in cities
+                                         for k in cities), cat='Binary')
+#1 if city i is assigned to hub k, 0 otherwise
 
-e = p.LpVariable.dicts('hasEdge', ((i,j) for i in cities
-                                         for j in cities), cat='Binary')
-
-x = p.LpVariable.dicts('Helper1', ((i,j) for i in cities
-                                         for j in cities), cat='Binary')
-
-y = p.LpVariable.dicts('Helper2', ((i,j,k,l) for i in cities
+y = p.LpVariable.dicts('Flow', ((i,j,k,l) for i in cities
                                           for j in cities
                                           for k in cities
                                           for l in cities), cat='Binary')
+#1 if i->k->l->j is the path from i to j, 0 otherwise
 
-a = p.LpVariable.dicts('Helper4', ((i,j) for i in cities
-                                         for j in cities), cat='Binary')
 #Objective function
 
 Parcel += p.lpSum(flow[i][j]*(col*cost[i][k]+
-                           trns*cost[k][l]+dist*cost[l][j])
+                           trns*cost[k][l]+dist*cost[l][j])*y[(i,j,k,l)]
                for i in cities
                for j in cities
                for k in cities
-               for l in cities) + p.lpSum(hub_cost[i]*h[i] for i in cities)
-#col*cost[i][k]*e[(i,k)] and (trns*cost[k][l]+dist*cost[l][j])*e[(l,j)] are not
-#dependent on eachother. So the first part gets added whenever e[(i,k)]=1 and
-#the second part whenever e[(l,j)] = 1. While it is only supposed to happen
-#when both of them are 1.
+               for l in cities) + p.lpSum(hub_cost[k]*e[k,k] for k in cities)
+
 
 #Constraints
 for i in cities:
-    Parcel += e[(i,i)] - h[i] == 0
-    #Hubs are connected to themselves, cities are not
+    Parcel += p.lpSum(e[(i,k)] for k in cities) == 1
[…]

### 2021-06-03 ILP/ILP.tex
Update ILP.tex

\documentclass[11pt]{article} % use larger type; default would be 10pt
\usepackage{amsmath}
\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)

\usepackage{geometry} % to change the page dimensions
\geometry{a4paper} % or letterpaper (US) or a5paper or....

\title{ILP of Parcel Delivery}
\author{Bart Stolk}

\begin{document}
\maketitle

\subsection*{Input}
\begin{align*}
c_{ij} &= \text{cost between city i and j}  &i,j \in N\\
w_{ij} &= \text{flow between city i and j} &i,j \in N\\
f_{i} &= \text{cost to turn city i into a hub} &i \in N\\
\chi &= \text{collection cost multiplier} \\
\alpha &= \text{transfer cost multiplier} \\
\delta &= \text{distribution cost multiplier} \\
\end{align*}

\subsection*{Variables}
\begin{align*}
e_{ik} &=
\begin{cases} 
1, & \text{if city i is assigned to hub k} \quad i=1,2,\ldots,N \quad j=1,2,\ldots,N  \\ 
0, & \text{otherwise} 
\end{cases} \\
y_{ijkl} &=
\begin{cases} 
1, & \text{if there is a path } i \rightarrow k \rightarrow l \rightarrow j \quad i=1,2,\ldots,N \quad j=1,2,\ldots,N  \\ 
0, & \text{otherwise} 
\end{cases} \\
\end{align*}

\subsection*{Objective function}
$$min\,z=\sum^N_{i=1}\sum^N_{j=1}\sum^N_{k=1}\sum^N_{l=1}(w_{ij}(\chi c_{ik}+\alpha c_{kl}+\delta c_{lj})y_{ijlk} + \sum^N_{k=1}f_{k}e_{kk}$$

\subsection*{Constraints}
\begin{align*}
\sum^N_{k=1}e_{ik} &= 1 &i \in N \\
e_{ik} & \leq e_{kk} &i,k \in N \\
y_{ijkl} &\leq e_{ik} &i,j,k,l \in N \\
y_{ijkl} &\leq e_{jl} &i,j,k,l \in N \\
y_{ijkl} &\geq e_{ik} + e_{jl} - 1 &i,j,k,l \in N \\
\sum^N_{k=1}f_k e_{kk} &\geq 1
\end{align*}
\end{document}

### 2021-06-08 Minutes/Minutes_Modelleren_4.8.tex
Updates from Overleaf

\documentclass{article}
\usepackage[utf8]{inputenc}

\begin{document}
\section*{Minutes Wednesday}
Aanwezig: Bart, Dyami, Severijn
Voorzitter: Dyami
Secretaris: Bart

\subsection*{Discussed topics}
\begin{itemize}
    \item Incorporating the small data set into the report.
    \item Finalizing the ILP implementation for the expanded problem.
    \item Giving feedback on the draft report of group 5.
    \item Rewriting the ILP definition in the report, put the entire thing at the top
    \item More graphics, pictures and the sorts.
    \item {Pitch, Who and about what}
\end{itemize}

\subsection*{Decisions}
\begin{itemize}
    \item Continue work on the small data set and the ILP implementation after meeting.
    \item Today (Tuesday), we will independently give feedback on portions of the draft report of group 5 and then come together to give feedback.
    \item Severijn is going to present the pitch.
\end{itemize}

\subsection*{Planning}
\begin{itemize}
    \item First draft Pitch done before Thursday.
    \item Finish ILP implementation before Thursday.
\end{itemize}
\end{document}

### 2021-06-08 Reports/draft_report.tex

@@
 \usepackage{graphicx}
+\usepackage{amsmath, amsfonts, amssymb}
+\usepackage{tikz}
+\usepackage{hyperref}
+\usetikzlibrary{positioning}
 \usepackage[table]{xcolor}
+\usepackage{listings}
 
@@
 
+%maak een foto foor op de titlepagina
+
+\newpage
+\section*{list of variables}
+\addcontentsline{toc}{section}{List of variables}
+% nog aanvullen
+
 \newpage
@@
 \addcontentsline{toc}{section}{Summary}
+In this project we explored the world of parcel delivery.
+Based on a dataset provided we explored various option to optimize the cost of delivering a parcel, this cost optimization was done by building or not building hubs near certain cities.
+The first algorithm applied was an intuitive algorithm it does not guarantee the optimal solution, but it does give a reasonable solution.
+The other algorithm applied was an ILP which is takes a lot longer to run but does give the best possible solution.
+
+%mag wat langer // moet opzichzelf het gehele verslag omvatten // alle belangrijke info en conlcusie
 
@@
 An important restraint in this matter is that packages can only move between hubs and cities and in particular not directly between cities themselves.
+%B: What about flow?
+// vertellen wat er in de data set staat
 
-Since this type of optimization problem is a common practice for companies, there already exists an ILP solver which can give the exact hubs for which have the lowest total cost, However, it has shown that these solvers can take up and lot op time and processing power.
+Since this type of optimization problem is a common practice for companies, there already exists an ILP solver which can give the exact hubs for which have the lowest total cost, However, it has shown that these solvers can take up and lot of time and processing power.
 Therefore it is preferable to get an idea for which routes and hubs will give the best, or at least a proper solution to the problem.
@@
 \section{Intuitive Algorithm}
-As mentioned in the introduction, it can be useful to use an intuitive algorithm in stead of a direct solver for the problem.
+\label{Inuitive algorithms}
+As mentioned in the introduction, it can be useful to use an intuitive algorithm instead of a direct solver for the problem.
 With a self made script it is possible to get a feel for the cities in which building a hub can have a major impact on reducing the cost of the project.
@@
 Intuitively this made sense as it probably the closest.
-If we take for example the Netherlands this would mean that a packet send from Groningen would be picked up by a hub close to Groningen and not by a hub in for example Maastricht.
-So the first step in the cost function algorithm was to go trough the list of cities and determine which hub from the list of inputted hubs was cheapest to collect from for each city.
-When this connection is determined the sum of the list multiplied by the collection factor is added to the cost and the the amount of packages are added to the packages of the hub city.
-This connection between the city and collection hub was stored in memory as it would be used later to determine transfers and delivery.
-As soon as this first step is complete we end up with a Dataframe with as columns the hub cities and rows the packages that need to go to each city.
-The next step is to go row by row and look up what connection that city had with which hub and transfer all the packages of that row to the correct hub again multiplying the amount times the correct cost associated with transfering from hub to hub.
+If we take for example the Netherlands this would mean that a package sent from Groningen would be picked up by a hub close to Groningen and not by a hub in for example Maastricht.
+So the first step in the cost function algorithm was to go trough the list of cities and determine which hub from the list of input hubs was cheapest to collect from for each city.
+When this connection is determined the sum of the list multiplied by the collection factor and is added to the cost and the the amount of packages are added to the packages of the hub city.
+%B: Very long, difficult to follow sentence.
+Split up in 2 maybe?
+
+        The connection between the city and collection hub was stored in memory as it would be used later to determine transfers and delivery.
+As soon as this first step is complete we end up with a Data frame with as columns the hub cities and rows the packages that need to go to each city.
+The next step is to go row by row and look up what connection that city had with which hub and transfer all the packages of that row to the correct hub again multiplying the amount times the correct cost associated with transferring from hub to hub.
 Then the total of that row at the correct hub will be delivered to the city and thus multiplying the total number of packages by the cost to deliver from that hub to the city and the delivery factor.
@@
 Now we only need to add the cost of building the hubs and we have the total cost of using specific hubs.
+%iets compacter wellicht
 
@@
 The total process is divided into the collection from the cities to the hubs, followed by the transfer between hubs themselves, and as last the distribution of packages from the hubs to the desired cities.
-
+    % we moeten misschien nog even de namen van 2.3 2.5 enzo aanpassen aangezien het originele idee van het opsplitsen van collection, transer en distribution  niet meer van toepassing is
     \subsection{Collection of packages}  
-
-In this process the cost function will only give an output of the cost that is formed by combining the cost for collection packages.
+    \label{Collection of packages}
+    In this process the cost function will only give an output of the cost that is formed by combining the cost for collecting packages.
 To optimize the costs a python script is made such that, when a city is given to be a hub, the code calculates the cost for this specific hub and the cost if the city next in line would be a hub as well.
@@
 If the cost rises, the city will be discarded from the process.
-If this process is repeated with each time a different starting, a list will appear of possible combination of hubs and there costs.
+If this process is repeated with each time a different starting node, a list will appear of possible combination of hubs and their costs.
 To illustrate this process the outputs will be stated below if it were to start with city number 3 as a hub.
@@
 
-\begin{table}
+        \subsubsection{Result of the large data set}
+
+            Using the large data set given with 14 different cities, the results are stated below.
+
+            \begin{table}[h!]
+                \begin{center}
                 \begin{tabular}{|p{3cm}|p{3cm}|p{3cm}|}
@@
                     7 & 2& 3 \\
-
                     \hline
@@
                 \end{tabular}
-\caption{Table of the amount of times each city is used as a hub in every combination}
+                \end{center}
+                \caption{Table of the amount of times each city is used as a hub in every combination using the large data set}
                 \label{amount of times hubs are used in collection} 
@@
 
-\begin{table}
+            \begin{table}[h!]
+                \begin{center}
                     \begin{tabular}{ |p{3cm}|p{3cm}|p{4cm}|}
@@
                     \end{tabular}
+                \end{center}
             \caption{Table of the amount of times each city is used as a hub in every combination}
@@
 
-\subsection{Transfer between hubs}
+        \subsubsection{Result of the small data set}
+            For the small data set the same method is used as for the large file.
+First the amount of times a cities is added as a hub is counted.
+Thereafter, the result we be run again by the algorithm with the most common cities as hubs to create an optimal solution.
+
+             \begin{table}[h!]
+                \begin{center}
+                \begin{tabular}{|p{3cm}|p{3cm}|p{3cm}|}
+                    \hline
+                    \multicolumn{3}{|c|}{Amount of times cities are used as hubs} \\
+                    \hline
+                    City used as hub & Using only collection cost & Using total cost\\
+                    \hline
+                    1 &  8 & 5 \\
+                    \hline
+                    2 & 4   & 11 \\
+                    \hline
+                    3& 7 & 9\\
+                    \hline
+                    4  & 1 & 1 \\
+                    \hline
+                    5 & 8 & 9\\
+                    \hline
+                    6 & 1 & 8   \\
+                    \hline
+                    7 & 2& 3 \\
+                    \hline
+                    8 & 2 & 2 \\
+                    \hline
+                    9 &  7& 10 \\
+                    \hline
+                    10 & 11 & 11 \\
+                    \hline
+                \end{tabular}
+                \end{center}
+                \caption{Table of the amount of times each city is used as a hub in every combination with the small data set}
+                \label{amount of times hubs are used in collection} 
+            \end{table}
+
+    \subsection{Transfer between hubs} \label{sssec:num1}
         Transport to the same node is free if the node is a hub.
-To optimize the total cost an algorithm is made that will be looking at the node with the most flow of parcels to itself and make a hub of this node.
+To optimize the total cost an algorithm is made that will look at the node with the most flow of parcels to itself and makes a hub of this node.
 After calculating the cost, the next hub with the most flow of parcels to itself will be become a hub.
@@
 If there are two nodes with the same amount of parcels to itself the algorithm will choose the node with the most incoming parcels.
-Each node will be connected to the hub with the least cost to transport it there.
-For example if it cost 11 dollars to transport from node 11 to node 14 and 27 dollars to transport it to node 2, node 11 will be connected with hub 14.
+Each node will be connected to the hub with the least cost to transport it there, for example if it cost 11 dollars to transport from node 11 to hub 14 and 27 dollars to transport it to hub 2, node 11 will be connected with hub 14.
 This process is repeated until the algorithm goes through all nodes.
-To illustrate we will use the Large data assignment.
-
-\begin{table}[h]
+To illustrate we will use the Large and small data assignment.
+\subsubsection{Result of the large data set}
+        \begin{table}[h!]
             \begin{tabular}{|p{3cm}|p{3cm}|p{3cm}|}
                 \hline
-nodes that are hubs & flow of parcels to itself (last node) & Total cost\\
+                nodes that are hubs & flow of parcels to itself (last node) & Total cost (€ x 10^3)\\
                 \hline
-[14] &  181 & 182725 \\
+                [14] &  181 & 182.7 \\
                 \hline
-[14,2]   & 39 & 163586 \\
+                [14,2]   & 39 & 163.5 \\
                 \hline
-[14,2,15] & 26 & 169342 \\
+                [14,2,15] & 26 & 169.3 \\
                 \hline
-[14,2,15,6]  & 18 & 156535 \\
+                [14,2,15,6]  & 18 & 156.5 \\
                 \hline
-[14,2,15,6,13] & 18 & 161521\\
+                [14,2,15,6,13] & 18 & 161.5\\
                 \hline
-[14,2,15,6,13,4] & 11 & 167037   \\
+                [14,2,15,6,13,4] & 11 & 167.0   \\
                 \hline
-[14,2,15,6,13,4,3] & 9 & 177126   \\
+                [14,2,15,6,13,4,3] & 9 & 177.1   \\
                 \hline
-[14,2,15,6,13,4,3,9] & 9 & 187710    \\
+                [14,2,15,6,13,4,3,9] & 9 & 187.7    \\
+                \hline
             \end{tabular}
+        \caption{Table of nodes that were used as hubs and the associated costs}
         \end{table}
@@
 The cheapest situation here is where there are 4 hubs in nodes 2,6,14 and 15.
-We connect each node with the hub where the cost to transport a parcel is the lowest, the total cost in this situation would be 156 thousand dollars.
+We connect each node with the hub where the cost to transport a parcel is the lowest, the total cost in this situation would be €156.5 thousand.
 One can easily see that it is very profitable to make a hub out of node 14, as it is the node with the most incoming parcels (610).
 In the following subsections we will be looking at the results of the other algorithms and combine those to see if there are similarities.
+\subsubsection{Result of the small data set}
+(The results of the small data set are still being worked on, though this will be done identically to the large data set.)
+
+\subsection{Distribution of packages}
+\label{Distribution}
+The idea of minimising the cost of distributing packages is similar to that of collecting packages.
+However, a somewhat different algorithm was used to do so.
+First, the algorithm iterates over the cities.
+For each city, it calculates the cost of having that city as a hub using the cost function.
+After going through all the cities, the algorithm picks the city which generates the lowest cost and `permanently' adds this city to the list of hubs.
+Then the algorithm repeats the first few steps, iterating over the cities (excluding the one that has been added as hub already) and calculate which of those costs the least to add.
+This results in a list of 15 results for the large data set and 10 for the small data set, with each with of the results having one more hub than the one before it.
+
+It is the expectation that the first few results will result in lower costs.
+But, once you start adding `too many' hubs, the costs will start rising again.
+Though this is not assumed to be necessarily true, so the algorithm keeps running until all of the cities have been added as hubs.
+
+\subsubsection{Result of the large data set}
+First, the algorithm was run to minimise distribution costs for the large data set.
+In table \ref{DistAlg}.
+To prevent repeating the entire list of hubs, the `Chosen hub' column will only include the hub that was added in that iteration of the algorithm.
+The costs have also been formulated in thousands rounded to 1 decimal for readability.
+
+\begin{table}[h]
+\centering
+\begin{tabular}{||c|c|c||}
+\hline
+    Chosen hub & Total distribution costs (€ x 10^3)& Total costs (€ x 10^3)  \\
+\hline
+\hline
+    10 & 64.2 & 169.2 \\
+    \hline
+    2 & 48.4 & 166.2 \\
+    \hline
+    14 & 35.6 & 153.3 \\
+    \hline
+    8 & 27.7 & 153.0 \\
+    \hline
+    9 & 22.0 & 159.6 \\
+    \hline
+    4 & 17.2 & 168.3 \\
+    \hline
+    6 & 13.9 & 173.3 \\
+    \hline
+    3 & 11.0 & 183.2 \\
+    \hline
+    1 & 8.2 & 192.2 \\
+    \hline
+    13 & 6.3 & 200.3 \\
+    \hline
+    12 & 4.8 & 208.0 \\
+    \hline
+    5 & 2.0 & 217.1 \\
+    \hline
+    11 & 0.9 & 230.2 \\
+    \hline
+    15 & 0 & 241.0 \\
+    \hline
+\end{tabular}
+\caption{\centering Order in which the cities got added as hubs by the algorithm based on distribution costs}
+\label{DistAlg}
+\end{table}
+
+As can be seen, the distribution costs decrease with each added hub until they reach 0.
+This is a result of distribution costs from hubs to themselves costs 0.
+So more of the actual costs are made up of transfer costs as more hubs get added.
+However, the total costs, displayed in the last column, have quite different results.
+At first, the costs decrease.
+But, after adding hub 9 the total costs increase from there onward.
+So, based on these results, a solution would be to pick hubs 2, 8, 10 and 14 with a total cost of 152974.
+
+The algorithm was also used based on total costs to compare how much the results differed from basing it on solely distribution costs, the results of which can be found in table \ref{DistAlg2}.
+With this, the hubs 6, 10 and 14 form an optimal solution.
+Interestingly, the order in which certain hubs appear in this version is somewhat similar to when it was based on just distribution costs.
+For example, the first 5 hubs to be added in both cases are the same, namely 2, 6, 8, 10 and 14.
+Even though their order is different.
+From this we might conclude that these cities are attractive candidates to be hubs in general.
+
+\begin{table}[h]
+\centering
+\begin{tabular}{||c|c||}
+\hline
+    Chosen hub & Total costs (€ x 10^3) \\
+\hline
+\hline
+    10 & 169.3 \\
+    \hline
+    14 & 156.3 \\
+    \hline
+    6 & 150.4 \\
+    \hline
+    8 & 153.2 \\
+    \hline
+    2 & 156.9 \\
+    \hline
+    5 & 161.6 \\
+    \hline
+    13 & 169.0 \\
+    \hline
+    12 & 178.7 \\
+    \hline
+    4 & 185.3 \\
+    \hline
+    1 & 194.3 \\
+    \hline
+    3 & 204.3 \\
+    \hline
+    7 & 216.4 \\
+    \hline
+    9 & 229.2 \\
+    \hline
+    11 & 242.3 \\
+    \hline
+    15 & 247.2 \\
+    \hline
+\end{tabular}
+
+\caption{\centering Order in which the cities got added as hubs by the algorithm based on total costs}
+\label{DistAlg2}
+\end{table}
+\subsubsection{Result of the small data set}
+(The results of the small data set are still being worked on, though this will be done identically to the large data set.)
 
@@
     \subsection{Result of the combined algorithms}
-After combining and analysing the results of the different algorithms, a hub will be opened at nodes 14 and 6 as these nodes were also hubs in the cheapest situation of all different algorithms that were used.
+    \subsubsection{Result of the large data set}
+By combining and analysing the results of the different algorithms, a hub will be opened at nodes 14 and 6 as these nodes were also hubs in the cheapest situation of all different algorithms that were used.
 By opening a hub at node 14 the cost for collecting , transferring and distributing can be reduced, as it is the node with the most amount of incoming parcels.
-The total cost in this situation where we only have two hubs is 144870.
-Our intuitive algorithm will then go through the nodes one by one and check which combination (with node 14 and 6) is the cheapest, this node will then be added to the list of hubs.
-This process is repeated until we go through all nodes.
+The total cost in this situation where we only have two hubs is €144.870.
+The intuitive algorithm will then go through the nodes one by one and check which combination (with node 14 and 6) is the cheapest, this node will then be added to the list of hubs.
+This process is repeated until the algorithm goes through all nodes.
 
@@
 \hline
-Nodes that are hubs & Total cost\\
+Nodes that are hubs & Total cost (€ x 10^3)\\
 \hline
-[6,14] &  144870 \\
+[6,14] &  144.8 \\
 \hline
-[6,8,14]   & 144532\\
+[6,8,14]   & 144.5\\
 \hline
-[2,6,8,14] & 148189 \\
+[2,6,8,14] & 148.1 \\
 \hline
-[2,5,6,8,14]  & 152593  \\
+[2,5,6,8,14]  & 152.5  \\
 \hline
-[2,5,6,8,12,14] & 160386 \\
+[2,5,6,8,12,14] & 160.3 \\
+\hline
 \end{tabular}
+\caption{Table of nodes that were used as hubs in the initiative algorithm}
 \end{table}
@@
 The most advantageous situation that was found with the intuitive algorithm is 
-where we make 3 hubs out of nodes 6,8 and 14.
-By doing this the total cost is reduced to 144532 dollars which is 338 dollars cheaper than the previous one.
-(heb dit snel gemaakt verander het als je wilt)
+where we make 3 hubs out of nodes 6,8 and 14. By doing this the total cost is reduced to €144.532 which is €338 cheaper than the previous one.
+
+\begin{figure}[h]
+\centering
+\begin{tikzpicture}[main/.style = {draw, circle,minimum size=0.8cm}]
+\node[main] (6) {6};
+\node[main] (8) [below right=1.5cm of 6] {8};
+\node[main] (14) [below left=1.5cm of 6] {14};
+
+\node[main] (1) [above=of 6,xshift=-1.7cm,yshift=-1.7cm] {1};
+\node[main] (2) [above=of 6,xshift=-1.3cm,yshift=-0.9cm] {2};
+\node[main] (3) [above=of 6,xshift=-0.5cm,yshift=-0.5cm] {3};
+\node[main] (5) [above=of 6,xshift=0.5cm,yshift=-0.5cm] {5};
+\node[main] (7) [above=of 6,xshift=1.3cm,yshift=-0.9cm] {7};
+
+\node[main] (4) [right=of 8] {4};
+\node[main] (11) [below right=of 8] {11};
+\node[main] (12) [below=of 8] {12};
+
+\node[main] (9) [left=of 14] {9};
+\node[main] (13) [below left=of 14] {13};
+\node[main] (15) [below=of 14] {15};
+\node[main] (10) [below right=of 14] {10};
+
+\draw[-] (8) -- (6);
+\draw[-] (14) -- (6);
+\draw[-] (8) -- (14);
+\draw[-] (10) -- (14);
+
+\draw[-] (1) -- (6);
+\draw[-] (2) -- (6);
+\draw[-] (3) -- (6);
+\draw[-] (5) -- (6);
+\draw[-] (7) -- (6);
+
+\draw[-] (4) -- (8);
+\draw[-] (11) -- (8);
+\draw[-] (12) -- (8);
+
+\draw[-] (9) -- (14);
+\draw[-] (13) -- (14);
+\draw[-] (15) -- (14);
+\end{tikzpicture}
+\caption{A graph of the optimal solution}
+\end{figure}
+
+\subsubsection{Result of the small data set}
+(The results of the small data set are still being worked on, though this will be done identically to the large data set.)
 
@@
 \section{The ILP solver}
-We are given a set N of nodes where we can locate a hub.
-The cost of opening a hub at node i $\in$ N is $F_i$ and the cost of transporting a parcel to node j from node i is $c_{i_j}$, i $\in$ N, j $\in$ N.
-We need to decide on which nodes we shall open a hub, and from which hub the nodes will be connected.
-The goal is to minimize the total cost.
+\label{LPSolver}
+With the intuitive algorithm now in place, an ILP solver for the problem.
+In the first subsection it will be detailed how the ILP was modelled.
+This includes the various inputs, the chosen variables, the objective function and the constraints.
+Then, the results of the ILP are discussed by comparing it to the results of the intuitive algorithm.
+The full ILP definition and the Python program used to solve it can be found in Appendix \ref{ILPDef} and \ref{ILPPyth}.
+
 \subsection{ILP implementation}
-    \subsection{IlP result}
+For the implementation of the ILP the actual values of the inputs will not be used, rather variables to represent any valid input for the ILP.
+The actual values can be found in <add ref>.
+A network of n nodes given by the set N represent the city.
+A city i $\in$ N can be transformed into a hub with cost $f_i$.
+The amount of packages sent, or flow, from a city i $\in$ N and a city j $\in$ N is given by $w_{ij}$.
+The cost to send one unit of flow from a city i $\in$ N to a city j $\in$ N is given by $c_{ij}$.
+It is important to note that the costs are symmetric.
+That is, $c_{ij} = c_{ji}$ for all i,j $\in$ N.
+But the flow is not necessarily symmetric.
+Lastly, there are multipliers for collecting (from city to hub), transfer (from hub to hub) and distribution (from hub to city).
+These are given the names $\chi$, $\alpha$, $\delta$ respectively.
 
+Given these inputs, it is now the objective to minimalise the costs of sending all the packages by choosing the right hubs and connections between cities and hubs.
+We start by introducing a binary variable $e_{ik}$ with i,k $\in$ N as follows:
+
+$$
+e_{ik} =
+\begin{cases} 
+1, & \text{if city i is assigned to hub k} \quad i,j \in N  \\ 
+0, & \text{otherwise}
+\end{cases} \\
+$$
+
+Note that $e_{ik}$ does not consider any connections between hubs.
+This is not needed however, as the connections between hubs are `implied' in the objective function.
+Additionally, with this variable we can also define what constitutes a hub in this ILP; namely a city k is a hub whenever $e_{kk} = 1$.
+
+Next, a means is required that indicates which path is valid for a package to take from a city i to j.
+A way to do this is by requiring that, for some i,j,k,l \in N, both $e_{ik} = 1$ and $e_{lj} = 1$.
+However, this would not be linear.
+Instead we introduce a helper variable $y_{ijkl}$ to mean the same as that:
+
+$$
+y_{ijkl} =
+\begin{cases} 
+1, & \text{if there is a path } i \rightarrow k \rightarrow l \rightarrow j \quad i,j,k,l \in N \\ 
+0, & \text{otherwise} 
+\end{cases}
+$$
+
+To make this variable mean the same as what is desired, the following three constraints are introduced:
+
+\begin{align*}
+    y_{ijkl} &\leq e_{ik} \\
+    y_{ijkl} &\leq e_{jl} \\
+    y_{ijkl} &\geq e_{ik} + e_{jl} - 1
+\end{align*}
+
+The validity of these constraints is shown in table \ref{tab:Helper}.
+
+\begin{table}[h]
+
+\centering
+\begin{tabular}{|p{1cm}|p{1cm}||p{2cm}|p{1cm}|}
+\hline
+$e_{ik}$ & $e_{jl} $ & constraints & $y_{ijkl}$\\
+\hline
+0 & 0 & 
+  \begin{tabular}{@{}c@{}c@{}c@{}}
+  y_{ijkl} &\leq 0 \\ 
+  y_{ijkl} &\leq 0 \\
+  y_{ijkl} &\geq -1
+  \end{tabular}
+& 0 \\
+\hline
+0 & 1 & 
+  \begin{tabular}{@{}c@{}c@{}c@{}}
+  y_{ijkl} &\leq 0 \\ 
+  y_{ijkl} &\leq 1 \\
+  y_{ijkl} &\geq 0
+  \end{tabular}
+& 0 \\
+\hline
+1 & 0 & 
+  \begin{tabular}{@{}c@{}c@{}c@{}}
+  y_{ijkl} &\leq 1 \\ 
+  y_{ijkl} &\leq 0 \\
+  y_{ijkl} &\geq 0
+  \end{tabular}
+& 0 \\
+\hline
+1 & 1 & 
+  \begin{tabular}{@{}c@{}c@{}c@{}}
+  y_{ijkl} &\leq 1 \\ 
+  y_{ijkl} &\leq 1 \\
+  y_{ijkl} &\geq 1
+  \end{tabular}
+& 1 \\
+\hline
+\end{tabular}
+\caption{Validity of $y_{ijkl}=e_{ik}e_{lj}$ for the given constraints} 
+\label{tab:Helper}
+\end{table}
+
+With this helper variable defined it is now possible to construct the objective function.
+First of all, the cost for a unit of flow sent along a certain path is needed.
+Given a path from $i \rightarrow k \rightarrow l \rightarrow j$ with i,j,k,l $\in$ N and the relevant multipliers we get that sending one unit of flow over this path is $\chi c_{ik} + \alpha c_{kl} + \delta c_{lj}$.
+However, we want to send more than just one package.
+So, it is multiplied by the total flow between the sending city i and receiving city j, namely $w_{ij}$.
+To make sure that the chosen path is valid, that is city i is assigned to hub k and city j is assigned to hub l, it is also multiplied by $y_{ijkl}$.
+Summing over i,j,k,l $\in$ N, the total costs for sending packages over the chosen hubs and connections is:
+$$\sum^N_{i=1}\sum^N_{j=1}\sum^N_{k=1}\sum^N_{l=1} w_{ij}(\chi c_{ik} + \alpha c_{kl} + \delta c_{lj})y_{ijkl}$$
+
+On top of that, there is also the costs of establishing hubs at certain cities.
+We multiply this cost for a certain city k $\in$ N by $e_{kk}$, since only hubs are connected to themselves.
+Summing over k, this gives the total cost of establishing the hubs as:
+$$\sum^N_{k=1} f_k e_{kk}$$
+
+Combining this with the previous part and the fact that the objective function needs to be minimised we obtain that the objective function is:
+$$min \sum^N_{i=1}\sum^N_{j=1}\sum^N_{k=1}\sum^N_{l=1} w_{ij}(\chi c_{ik} + \alpha c_{kl} + \delta c_{lj})y_{ijkl} + \sum^N_{k=1} f_k e_{kk}$$
+
+Lastly, some additional constraints are introduced.
+Firstly, since every city needs to be assigned to exactly one hub, $\sum^N_k e_{ik} = 1$ for all i $\in$ N is introduced as a constraint.
+Additionally, the constraint $e_{ik} \leq e_{kk}$ for all i,k $\in$ N is set as a constraint as well.
+For the latter, when $e_{kk} = 0$, it is a city.
+Thus, there can be no other cities assigned to it, which that constraints makes sure of.
+When $e_{kk} = 1$, it is a hub, meaning that there can be a different city assigned to it, though it is not required.
+Finally, a constraint is introduced that requires there to be at least one hub, namely $\sum^N_k e_{kk} \geq 1$.
+With these constraints the ILP definition is finalised and it can now be used to calculate an optimal solution.
+
+\subsection{ILP result}
+The ILP model was implemented into Python using PuLP, with CPLEX as solver.
+The relevant data was imported from an excel file using Pandas.
+The full program can be found in Appendix \ref{ILPPyth}.
+After running the program it returned an optimal solution of 144.369.
+Interestingly, the hubs chosen by this program were 6, 8 and 14, the exact same ones as the result from the intuitive algorithm.
+However, that resulted in a cost of 144.532, which is a difference of 163.
+
+Upon closer inspection, it was discovered that one connection differed from the intuitive algorithm.
+In the ILP optimal solution, a graph of which can be found in figure \ref{ILPSol}, city 10 was connected to hub 14 instead of hub 6.
+The connections between cities 10 and 14 and between 10 and 6 cost the same, namely 9.
+So the difference in cost is not due to collection or distribution costs, rather it is due to the transfer costs of sending packages through hub 14 instead of 6.
+The similarity between the two results does suggest that the intuitive algorithm is relatively accurate.
+
+\begin{figure}[h]
+
+\centering
+\begin{tikzpicture}[main/.style = {draw, circle,minimum size=0.8cm}]
+\node[main] (6) {6};
+\node[main] (8) [below right=1.5cm of 6] {8};
+\node[main] (14) [below left=1.5cm of 6] {14};
+
+\node[main] (1) [above=of 6,xshift=-1.7cm,yshift=-1.7cm] {1};
+\node[main] (2) [above=of 6,xshift=-1.3cm,yshift=-0.9cm] {2};
+\node[main] (3) [above=of 6,xshift=-0.5cm,yshift=-0.5cm] {3};
+\node[main] (5) [above=of 6,xshift=0.5cm,yshift=-0.5cm] {5};
+\node[main] (7) [above=of 6,xshift=1.3cm,yshift=-0.9cm] {7};
+\node[main] (10) [above=of 6,xshift=1.7cm,yshift=-1.7cm,draw=green!80,fill=green!20] {10};
+
+\node[main] (4) [right=of 8] {4};
+\node[main] (11) [below right=of 8] {11};
+\node[main] (12) [below=of 8] {12};
+
+\node[main] (9) [left=of 14] {9};
+\node[main] (13) [below left=of 14] {13};
+\node[main] (15) [below=of 14] {15};
+\node[main] (10-2) [below right=of 14,draw=red!80,fill=red!20] {10};
+
+\draw[-] (8) -- (6);
+\draw[-] (14) -- (6);
+\draw[-] (8) -- (14);
+
+\draw[-] (1) -- (6);
+\draw[-] (2) -- (6);
+\draw[-] (3) -- (6);
+\draw[-] (5) -- (6);
+\draw[-] (7) -- (6);
+\draw[-,draw=green!80] (10) -- (6);
+
+\draw[-] (4) -- (8);
+\draw[-] (11) -- (8);
+\draw[-] (12) -- (8);
+
+\draw[-] (9) -- (14);
+\draw[-] (13) -- (14);
+\draw[-] (15) -- (14);
+\draw[-,draw=red!80] (10-2) -- (14);
+\end{tikzpicture}
+\caption{A graph of the optimal ILP solution. The 10 coloured red indicates the result of the intuitive algoritm while the 10 coloured green indicates the result of the ILP solution.}
+\label{ILPSol}
+\end{figure}
 \newpage
+
 \section{Expanded algorithm}
@@
 To make this problem more realistic a situation is made where all the packages are collected and distributed using a bus.
-In this particular case, a bus can hold up to and including a total of one hundred packages.
+In this particular case, a bus can hold up to and including a total of one hundred packages~\cite{Packages per bus}.
 If more packages are needed to move every package, the delivery company will have to send two or more busses.
+Between the hubs, the company is able to use cargo trucks, such that everything can be send in one vehicle.
+This scenario is translated into the algorithm by letting all factors of moving packages to 1 and multiplying this by the amount of busses that are necessary.
 
-This scenario is translated into the algorithm by letting all factors of moving packages
+    \subsection{Intuitive algorithm result}
+
+        By evaluating the expanded intuitive algorithm a separation is made between the different intuitive algorithms used in Chapter~\ref{Inuitive algorithms}.
+The expansion of the model will likewise as the standard data have a result which is a combination of the multiple intuitive algorithms.
+
+        \subsubsection{Algorithm used for the collection costs}
+            For the expanded algorithm the same method will be used as in chapter~\ref{Collection of packages}.
+By adding a city as hub one by one and adding the city as a hub when the costs are lowered, the program is able to make a list of how many times cities are used as hub.
+With this the cities that are overall valuable to transform into a hub will stand out.
+
+             \begin{table}[h!]
+                \begin{center}
+                \begin{tabular}{|p{3cm}|p{3cm}|p{3cm}|}
+                    \hline
+                    \multicolumn{3}{|c|}{Amount of times cities are used as hubs} \\
+                    \hline
+                    City used as hub & Using only collection cost & Using total cost\\
+                    \hline
+                    1 &  5 & 4 \\
+                    \hline
+                    2 & 7   & 12 \\
+                    \hline
+                    3& 6 & 7\\
+                    \hline
+                    4  & 1 & 1 \\
+                    \hline
+                    5 & 7 & 7\\
+                    \hline
+                    6 & 1 & 10   \\
+                    \hline
+                    7 & 1& 2 \\
+                    \hline
+                    8 & 1 & 1 \\
+                    \hline
+                    9 &  10& 10 \\
+                    \hline
+                    10 & 11 & 12 \\
+                    \hline
+                    11 & 1& 1 \\
+                    \hline
+                    12 & 1 & 1 \\
+                    \hline
+                    13 & 2& 1 \\
+                    \hline
+                    14 & 14& 14 \\
+                    \hline
+                \end{tabular}
+                \end{center}
+                \caption{Table of the amount of times each city is used as a hub in every combination using the large data set for the expanded algorithm}
+                \label{amount of times hubs are used in collection expanded algorithm} 
+            \end{table}
+
+            Applying this data to the original algorithm the following hubs could be preferable.
+
+             \begin{table}[h!]
+                \begin{center}
+                    \begin{tabular}{ |p{3cm}|p{3cm}|p{4cm}|}
+                        \hline
+                        \multicolumn{3}{|c|}{Cost and combination of hubs with certain input} \\
+                        \hline
+                        Input into the algorithm & Output -  collection Cost (€ x 10^3) & Output - total Cost (€ x 10^3)\\
+                        \hline
+                        [14] &  78.3 | [14, 1, 3] & 153.4| [14, 1, 2, 3, 6]  \\
+                        \hline
+                        [10] & 76.1 | [10, 2, 14]   & 139.5 | [10, 2, 14] \\
+                        \hline
+                        [9,10 ]& 88.4 | [6, 10, 2, 14] & 148.8 |  [9, 10, 2, 14]\\
+                        \hline
+                        [6, 14] &  66.6 | [6, 14]  &  131.7 | [6, 14] \\
+                        \hline
+                        [6, 14, 8] & 75.6 | [2, 10, 14] & 136.3 | [6,14,8]  \\
+                        \hline
+                    \end{tabular}
+                \end{center}
+            \caption{Table of the amount of times each city is used as a hub in every combination}
+            \label{amount of times hubs are used in collection} 
+            \end{table}
+
+            This table shows that most likely turning cities 6 and 14 into hubs will result into a low cost. Moreover, cities 2, 10 and 8 also have a high probability of ending up in the final result.
+
+        \subsubsection{Algorithm used for transfers between hubs}
+            The algorithm that was used in subsection \ref{sssec:num1} on page \pageref{sssec:num1} will be implemented for this situation (more realistic) where the packages are collected and distributed using a bus.
+We will illustrate this by using the large data assignment.
+
+            \begin{table}[h!]
+            \begin{tabular}{|p{3cm}|p{3cm}|p{3cm}|}
+            \hline
+            nodes that are hubs & flow of parcels to itself (last node) & Total cost (€ x 10^3) \\
+            \hline
+            [14] &  181 & 404.3 \\
+            \hline
+            [14,2]   & 39 & 282.4 \\
+            \hline
+            [14,2,15] & 26 & 273.1 \\
+            \hline
+            [14,2,15,6]  & 18 & 228.9 \\
+            \hline
+            [14,2,15,6,13] & 18 & 222.2\\
+            \hline
+            [14,2,15,6,13,4] & 11 & 213.6   \\
+            \hline
+            [14,2,15,6,13,4,3] & 9 & 214.9   \\
+            \hline
+            [14,2,15,6,13,4,3,9] & 9 & 218.5    \\
+            \hline
+            [14,2,15,6,13,4,3,9,7] & 9 & 222.9    \\
+            \hline
+            \end{tabular}
+            \caption{Table of nodes that were used as hubs and the associated costs}
+            \end{table}
+
+The cheapest situation here is where we have 6 hubs in nodes 2,4,6,13,14,15, the total cost in this situation is €213.621 which is €57086 more expensive than before.
+An explanation for this is that the packages are not sent on one go, but in two or more times, which would ultimately lead to more costs.
+There is also a big difference in the locations of the hubs.
+First there were 4 (2,6,14 and 15) hubs and now two extra hubs are being established.
+In the following chapters the other algorithms will be implemented and combined so that in the end an optimal situation arises.
+
+\subsubsection{Algorithm used for distribution}
+The algorithm used for distribution cost minimisation for the addition to the problem is identical to the one in section \ref{Distribution}.
+The only modification is to the cost function.
+Running the algorithm based on distribution costs, given the change, results in table \ref{tab:DistAlgExp}.
+Again, the first column represents which city got chosen to add to the list of hubs and the costs have been expressed in thousands rounded to one decimal for readability.
+
+\begin{table}[h!]
+    \centering
+    \begin{tabular}{||c|c|c||}
+    \hline
+        Chosen hub & Total distribution costs (€ x 10^3) & Total costs (€ x 10^3)  \\
+    \hline
+    \hline
+        14 & 81.2 & 165.4 \\
+        \hline
+        3 & 45.2 & 136.5 \\
+        \hline
+        8 & 37.1 & 141.3 \\
+        \hline
+        2 & 29.5 & 146.3 \\
+        \hline
+        9 & 22.3 & 154.1 \\
+        \hline
+        4 & 17.5 & 164.0 \\
+        \hline
+        7 & 13.8 & 169.3 \\
+        \hline
+        1 & 11.0 & 179.9 \\
+        \hline
+        10 & 8.7 & 192.1 \\
+        \hline
+        13 & 6.9 & 201.1 \\
+        \hline
+        6 & 5.0 & 211.4 \\
+        \hline
+        12 & 3.5 & 220.0 \\
+        \hline
+        5 & 2.0 & 230.1 \\
+        \hline
+        11 & 0.9 & 243.8 \\
+        \hline
+        15 & 0 & 253.1 \\
+        \hline
+    \end{tabular}
+    \caption{\centering Order in which the cities got added as hubs by the algorithm based on distribution costs with the expanded problem}
+    \label{tab:DistAlgExp}
+\end{table}
+This result differs quite a bit from the results of section \ref{Distribution}.
+There are some elements that remain the same, such as the distribution costs decreasing for each added hub and the total costs decreasing at first but increasing later.
+However, the order in which hubs gets added is quite different.
+In this case, an optimal solution would be turning cities 3 and 14 into hubs.
+
+Once again, the algorithm is repeated but instead with a focus on the total costs.
+The results can be seen in table \ref{tab:DistAlgExp2}.
+Notably, just as in section \ref{Distribution}, we see that cities 2, 6, 8 and 14 are first to be chosen, though 10 is one of the last to be added in this case.
+The optimal solution in this case also only consists of 2 hubs, namely 6 and 14.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{||c|c||}
+    \hline
+        Chosen hub & Total costs (€ x 10^3) \\
+    \hline
+    \hline
+        14 & 165.4 \\
+        \hline
+        6 & 131.7 \\
+        \hline
+        2 & 133.2 \\
+        \hline
+        8 & 137.6 \\
+        \hline
+        13 & 144.4 \\
+        \hline
+        12 & 153.2 \\
+        \hline
+        5 & 162.0 \\
+        \hline
+        11 & 172.2 \\
+        \hline
+        4 & 182.5 \\
+        \hline
+        1 & 193.0 \\
+        \hline
+        3 & 203.9 \\
+        \hline
+        7 & 216.6 \\
+        \hline
+        9 & 223.0 \\
+        \hline
+        10 & 243.8 \\
+        \hline
+        15 & 253.1 \\
+        \hline
+    \end{tabular}
+    \caption{\centering Order in which the cities got added as hubs by the algorithm based on total costs with the expanded problem}
+    \label{tab:DistAlgExp2}
+\end{table}
+
+\subsection{ILP result}
+(This part of the expanded algorithm is still being worked on)
+
+\subsection{Difference between the two methods}
+(This section is still being worked on)
 
@@
 
-\section{Appendix} 
+%we moeten hier de opdracht nog even uitleggen opnieuw // ook in het begin duidelijker verwoorden 
+Looking back at the report, the main goal is evaluated.
+The aim was to create an algorithm which can find a near to perfect or at least a good solution to the package delivery problem.
+This problem is based on the delivery and collection of packages send between different cities.
+For each of these cities the option arises to build a so called hub.
+It is only possible to send packages from and to hubs.
+Hereby it is important to notice that it is not allowed to send a package between two cities if neither of them is a hub.
+
+It is preferable to create this algorithm to get a feel for which hubs will give the lowest total cost for the package delivery company.
+This is due to the fact that there already is an ILP solver which, while it takes quite a long time to run, gives the exact best options for the placement of hubs.
+Out of the combination of the multiple intuitive algorithms it has shown that building hubs in cities 6, 8 and 14 will give one of the lowest possible amount of money.
+With this combination the total cost results into €144.522 euros.
+When compared to the exact result, the ILP solver gives the same cities as hubs, namely 6, 8 and 14, but reaches a lower overall cost.
+The solver changes the connection from city 10 to hub 14 instead of hub 6, which is used in the intuitive algorithms.
+
+In conclusion, the intuitive algorithm gave a result which was only €163 away form the ILP solver.
+Given that the intuitive algorithm takes a fraction of the calculating time that the solver uses, it is possible to conclude that the algorithm works extremely well for this specific problem.
+%stukje over resultaat expanded algorithm
+By adding another element into the data set, the problem becomes more realistic.
+this is done by looking at the way packages are moved around.
+In this case the restrictions of the problem have been modified such that the packages are moved by busses which can hold up to 100 packages.
+Even in this expanded algorithm, it shows that the intuitive part assesses the problem quite well.
+
+\newpage
+\section{Bibliography}
+\begin{thebibliography}{9}
+
+\bibitem{Packages per bus}
+(2016) \textit{Omroepwest}
+\url{omroepwest.nl/nieuws/3313393/Tachtig-tot-110-pakketten-per-bus-bezorgers-hebben-het-druk-vlak-voor-Kerstmis}
+
+\end{thebibliography}
+
+\newpage
+
+\appendix
+\section{ILP definition}
+\label{ILPDef}
+\subsection*{Input}
+\begin{align*}
+c_{ij} &= \text{cost between city i and j}  &i,j \in N\\
+w_{ij} &= \text{flow between city i and j} &i,j \in N\\
+f_{i} &= \text{cost to turn city i into a hub} &i \in N\\
+\chi &= \text{collection cost multiplier} \\
+\alpha &= \text{transfer cost multiplier} \\
+\delta &= \text{distribution cost multiplier} \\
+\end{align*}
+
+\subsection*{Variables}
+\begin{align*}
+e_{ik} &=
+\begin{cases} 
+1, & \text{if city i is assigned to hub k} \quad i=1,2,\ldots,N \quad j=1,2,\ldots,N  \\ 
+0, & \text{otherwise} 
+\end{cases} \\
+y_{ijkl} &=
+\begin{cases} 
+1, & \text{if there is a path } i \rightarrow k \rightarrow l \rightarrow j \quad i=1,2,\ldots,N \quad j=1,2,\ldots,N  \\ 
+0, & \text{otherwise} 
+\end{cases} \\
+\end{align*}
+
+\subsection*{Objective function}
+$$min\,z=\sum^N_{i=1}\sum^N_{j=1}\sum^N_{k=1}\sum^N_{l=1}(w_{ij}(\chi c_{ik}+\alpha c_{kl}+\delta c_{lj})y_{ijlk} + \sum^N_{k=1}f_{k}e_{kk}$$
+
+\subsection*{Constraints}
+\begin{align*}
+\sum^N_{k=1}e_{ik} &= 1 &i \in N \\
+e_{ik} & \leq e_{kk} &i,k \in N \\
+y_{ijkl} &\leq e_{ik} &i,j,k,l \in N \\
+y_{ijkl} &\leq e_{jl} &i,j,k,l \in N \\
+y_{ijkl} &\geq e_{ik} + e_{jl} - 1 &i,j,k,l \in N \\
+\sum^N_{k=1}f_k e_{kk} &\geq 1
+\end{align*}
+
+\newpage
+
+\section{ILP Python implementation}
+\label{ILPPyth}
+This version of the implementation is specifically for the large data set. The implementation for the small data set is identical apart from some of the input.
+\begin{verbatim}
+    import pandas as pd
+import pulp as p
+import numpy as np
+#Input
+col = 3
+trns = 1
+dist = 2
+cities = range(15)
+
+file = 'Large.xlsx'
+xls = pd.ExcelFile(file)
+
+flow = pd.read_excel(xls, 'w')
+flow = flow.values[:,1:]
+
+cost = pd.read_excel(xls, 'c')
+cost = cost.values[:,1:]
+
+hub_cost = pd.read_excel(xls, 'f')
+hub_cost = hub_cost.values[:,1]
+hub_cost = np.append([13530], hub_cost)
+
+o = dict()
+d = dict()
+for i in cities:
+    o[i] = p.lpSum(flow[i][j] for j in cities)
+    d[i] = p.lpSum(flow[j][i] for j in cities)
+
+#Creating the model
+Parcel = p.LpProblem('ParcelDelivery', p.LpMinimize)
+
+#Variables
+e = p.LpVariable.dicts('hasEdge', ((i,k) for i in cities
+                                         for k in cities), cat='Binary')
+#1 if city i is assigned to hub k, 0 otherwise
+
+y = p.LpVariable.dicts('Flow', ((i,j,k,l) for i in cities
+                                          for j in cities
+                                          for k in cities
+                                          for l in cities), cat='Binary')
+#1 if i->k->l->j is the path from i to j, 0 otherwise
+
+#Objective function
+
+Parcel += p.lpSum(flow[i][j]*(col*cost[i][k]+
+                           trns*cost[k][l]+dist*cost[l][j])*y[(i,j,k,l)]
+               for i in cities
+               for j in cities
+               for k in cities
+               for l in cities) + p.lpSum(hub_cost[k]*e[k,k] for k in cities)
+
+#Constraints
+for i in cities:
+    Parcel += p.lpSum(e[(i,k)] for k in cities) == 1
+#Each city i is connected to exactly one hub k.
+
+    for k in cities:
+        Parcel += e[(i,k)] <= e[(k,k)]
+#There can only be connections from i to k if k is connected to itself
+#(meaning k is a hub)
+
+        for j in cities:
+            for l in cities:
+                Parcel += y[(i,j,k,l)] <= e[(i,k)]
+                Parcel += y[(i,j,k,l)] <= e[(j,l)]https://www.overleaf.com/project/607ecb97092dbc483a588800
+                Parcel += y[(i,j,k,l)] >= e[(i,k)] + e[(j,l)] - 1
+#With these constraints: y[(i,j,k,l)] = e[(i,k)]*e[(j,l)]
+
+Parcel += p.lpSum(e[k,k] for k in cities) >= 1
+#There must be at least one hub
+
+solver = p.CPLEX_CMD()
+res = Parcel.solve(solver)
+print("status:", res)
+print("OPT:", p.value(Parcel.objective))
+\end{verbatim}
+
 \end{document}

### 2021-06-14 Intuitive Algoritm/cost_functie.py
Updated

@@ -3,10 +3,12 @@ import matplotlib.pyplot as plt
 import math
 collection_factor = 1
 transfer_factor = 1
-distribution_factor = 2
-bus_capactity = 50
+distribution_factor = 1
+bus_capactity = 100
 
-file_name ='Large.xlsx'
+file_name ='data_klein.xlsx'
+
+
 class parcel():
 	def __init__(self,file):
 		xls = pd.ExcelFile(file)
@@ -14,7 +16,11 @@ class parcel():
 		self.flow = self.flow.transpose()
 		self.package_cost = pd.read_excel(xls, 'c')
 		self.package_cost = self.package_cost.transpose()
-		self.hub_cost =[13530] + list(pd.read_excel(xls, 'f')[13530])
+		if file == 'Large.xlsx':
+			self.hub_cost =[13530] + list(pd.read_excel(xls, 'f')[13530])
+		else:
+			self.hub_cost =[4388] + list(pd.read_excel(xls, 'f')[4388])
+
 		self.cities = len(self.flow.columns)
 
 	def calculate_cost(self,hubs):
@@ -56,4 +62,4 @@ class parcel():
 
 
 test = parcel(file_name) 
-print(test.calculate_cost([1,3]))
+print(test.calculate_cost([6,14]))

### 2021-06-15 Intuitive Algoritm/cost_functie.py
Added comments

@@ -11,6 +11,7 @@ file_name ='data_klein.xlsx'
 
 class parcel():
 	def __init__(self,file):
+		#Here the data is imported from the excel file and stored and variables to be used later
 		xls = pd.ExcelFile(file)
 		self.flow = pd.read_excel(xls, 'w')
 		self.flow = self.flow.transpose()
@@ -24,35 +25,38 @@ class parcel():
 		self.cities = len(self.flow.columns)
 
 	def calculate_cost(self,hubs):
+		# hubs are given as input variable in the form of a list so eg. [1,2,5]
 		cost = 0
-		index = {}
+		index = {} 
 		hubs = sorted(hubs)
 		flow_copy = self.flow.copy(deep=True)
 		flow_copy = flow_copy.drop(index = 'Unnamed: 0')
 		cost_copy = self.package_cost.copy(deep=True)
 		cost_copy = cost_copy.take([hub for hub in hubs])
+		# The following loop calculates the cost of collection and building of the hubs. 
 		for i in range(self.cities):
 			if i+1 in hubs:
+				#If statement checks if the hub is in the inputed if yes then the cost to built this hub is added to the total cost
 				index[i] = i+1
 				cost+= self.hub_cost[i]
 				continue
 			nearest_hub = cost_copy.loc[cost_copy[i] > 0 , i].idxmin()
+			#the nearest hub is determined by finding the hub that has the lowest collection cost for that city 
 			cost += cost_copy[i][nearest_hub] * flow_copy[i].sum()*math.ceil(flow_copy[i].sum()/bus_capactity)*collection_factor
+			#The amount of packages multiplied by the cost of collection of the nearest hub multiplied by the collection factor is added to the cost 
 			nearest_hub = nearest_hub 
 			index[i] = nearest_hub
+			#The city is linked to the hub 
 			flow_copy[nearest_hub-1] = flow_copy[i]+flow_copy[nearest_hub-1]
 			flow_copy = flow_copy.drop(columns = i)
 		for i in range(self.cities):
-			# print('index = ' + str(i))
+			#in this loop the cost of transport and delivery is determined. 
 			flowss = list(flow_copy.iloc[i])
 			to_hub = index[i] - 1 
 			prices = list(cost_copy[to_hub])
 			total_packages = sum(flowss)
 			transfer_cost = sum([int(flowss[i])*int(prices[i]) for i in range(len(flowss))])
[…]

### 2021-06-17 Reports/draft_report.tex
Added text

@@
 The first algorithm applied was an intuitive algorithm it does not guarantee the optimal solution, but it does give a reasonable solution.
+The intuitive algorithm worked by iteratively finding the hub that would give the lowest cost and adding that to the list of hubs.
+So the program would find the hub that would minimize cost and add it to the list, for example 14 then it would go through all hubs again and find for example that 8 would minimize cost.
+This would repeat till adding an extra hub would not lower cost any more.
 The other algorithm applied was an ILP which is takes a lot longer to run but does give the best possible solution.
+
+Interestingly, the results of the ILP solver came very close to the results of the intuitive algorithm.
+Both the ILP and the intuitive algorithm gave as the optimal solution that cities 14,8,6 should be hubs for a total cost of 144.522 euros.
+The only difference between the result of the ILP and the Intuitive algorithm is that ILP connected city 10 with hub 6 whilst the intuitive algorithm connected that city with hub 14.
+This difference was only 163 euros.
+
+To extend the problem, we wanted to make the model more realistic.
+We looked at what if the amount of packages would be limited by the capacity of a transport bus.
+So instead of using multipliers which we used in the original problem we capped the maximum amount a bus could carry and determined the cost by the amount of transport busses necessary.
 
@@
     In this process the cost function will only give an output of the cost that is formed by combining the cost for collecting packages.
-To optimize the costs a python script is made such that, when a city is given to be a hub, the code calculates the cost for this specific hub and the cost if the city next in line would be a hub as well.
+To optimize the costs a Python script is made such that, when a city is given to be a hub, the code calculates the cost for this specific hub and the cost if the city next in line would be a hub as well.
 If the total cost drops by adding city as a hub, this city will be added to list of cities that are hubs.
@@
 If this process is repeated with each time a different starting node, a list will appear of possible combination of hubs and their costs.
-To illustrate this process the outputs will be stated below if it were to start with city number 3 as a hub.
+To illustrate this process the outputs will be stated below in table~\ref{Collection Algorithm} if it were to start with city number 3 as a hub.
 
@@
                 \hline
-                Cost of the combination(€ x 10^3) & combination on cities as hubs \\ [0.5ex] 
-                \hline\hline
+                Cost of the combination(€ x 10^3) & combination of cities as hubs \\ [0.5ex] 
+                \hline
+                \hline
                 169.3 & [3,1]  \\ 
@@
                 \begin{center}
-                \begin{tabular}{|p{3cm}|p{3cm}|p{3cm}|}
-                    \hline
-                    \multicolumn{3}{|c|}{Amount of times cities are used as hubs} \\
+                \begin{tabular}{||c|c|c||}
+                    \hline
+                    \multicolumn{3}{||c||}{Amount of times cities are used as hubs} \\
+                    \hline
                     \hline
                     City used as hub & Using only collection cost & Using total cost\\
+                    \hline
                     \hline
@@
 
-            When the algorithm is used for different input combinations based on ~\ref{amount of times hubs are used in collection} the outputs and results show different possible answers with supposedly relative low cost.
+            When the algorithm is used for different input combinations based on table ~\ref{amount of times hubs are used in collection} the outputs and results show different possible answers with supposedly relative low cost.
 Hereby the algorithm will add a city as hub if this results in a lower cost by the principle given in the beginning of the section.
@@
                 \begin{center}
-                    \begin{tabular}{ |p{3cm}|p{3cm}|p{4cm}|}
-                        \hline
-                        \multicolumn{3}{|c|}{Cost and combination of hubs with certain input} \\
+                    \begin{tabular}{||c|c|c||}
+                        \hline
+                        \multicolumn{3}{||c||}{Cost and combination of hubs with certain input} \\
+                        \hline
                         \hline
                         Input into the algorithm & Output -  collection Cost (€ x 10^3) & Output - total Cost (€ x 10^3)\\
+                        \hline
                         \hline
@@
                 \begin{center}
-                \begin{tabular}{|p{3cm}|p{3cm}|p{3cm}|}
-                    \hline
-                    \multicolumn{3}{|c|}{Amount of times cities are used as hubs} \\
+                \begin{tabular}{||c|c|c||}
+                    \hline
+                    \multicolumn{3}{||c||}{Amount of times cities are used as hubs} \\
                     \hline
@@
                     \hline
-                    1 &  8 & 5 \\
-                    \hline
-                    2 & 4   & 11 \\
-                    \hline
-                    3& 7 & 9\\
-                    \hline
-                    4  & 1 & 1 \\
-                    \hline
-                    5 & 8 & 9\\
-                    \hline
-                    6 & 1 & 8   \\
-                    \hline
-                    7 & 2& 3 \\
-                    \hline
-                    8 & 2 & 2 \\
-                    \hline
-                    9 &  7& 10 \\
-                    \hline
-                    10 & 11 & 11 \\
+                    1 & 6 & 6 \\
+                    \hline
+                    2 & 7   & 9\\
+                    \hline
+                    3& 5& 7\\
+                    \hline
+                    4  & 2 & 2 \\
+                    \hline
+                    5 & 4 & 7\\
+                    \hline
+                    6 & 4 & 8   \\
+                    \hline
+                    7 & 4& 1 \\
+                    \hline
+                    8 & 1 & 1 \\
+                    \hline
+                    9 &  1& 1 \\
                     \hline
@@
                 \caption{Table of the amount of times each city is used as a hub in every combination with the small data set}
-                \label{amount of times hubs are used in collection} 
-            \end{table}
+                \label{amount of times hubs are used in collection small} 
+            \end{table}
+
+            Using this data an estimation is made for the following hubs.
+
+            \begin{table}[h!]
+                \begin{center}
+                    \begin{tabular}{||c|c|c||}
+                        \hline
+                        \multicolumn{3}{||c||}{Cost and combination of hubs with certain input} \\
+                        \hline
+                        Input into the algorithm & Output -  collection Cost (€ x 10^3) & Output - total Cost (€ x 10^3)\\
+                        \hline
+                        [1] & 35.1| [1,2,3,6] & 52.0| [1, 2, 3, 5, 6]  \\
+                        \hline
+                        [2] & 34.5 | [2,3,5,7]   & 50.4 | [2,3,5,6] \\
+                        \hline
+                        [3]& 33.4 | [3,5,7] & 50.4 |  [2,3,5,6]\\
+                        \hline
+                        [5] &  35.9| [5,1,2,4,6]  &  52.0 | [5,1,2,3,6] \\
+                        \hline
+                        [7] &  31.5| [7,1] & 50.8 |  [1,2,5,7] \\
+                        \hline
+                        [6] & 26.3 | [6,2] & 46.3 |  [2,4,6]\\
+                        \hline
+                        [2, 5] & 33.1 | [2,5,4,6] & 50.4| [2, 5, 3,6]  \\
+                        \hline
+                    \end{tabular}
+                \end{center}
+            \caption{Table of the amount of times each city is used as a hub in every combination small data set}
+            \label{amount of times hubs are used in collection small} 
+            \end{table}
+
+            With the values out of table~\ref{amount of times hubs are used in collection small} it shows that making hubs in cities 2,4 and 6 will give an approximation of the optimal solution.
+This combination gives a total amount of 46330 euros.
 
     \subsection{Transfer between hubs} \label{sssec:num1}
-        Transport to the same node is free if the node is a hub.
-To optimize the total cost an algorithm is made that will look at the node with the most flow of parcels to itself and makes a hub of this node.
-After calculating the cost, the next hub with the most flow of parcels to itself will be become a hub.
-If the solution becomes cheaper by adding a hub, the node is added to a list of hubs.
-If the total cost increases, nothing happens and the next node (most parcels to itself) becomes a hub.
-If there are two nodes with the same amount of parcels to itself the algorithm will choose the node with the most incoming parcels.
-Each node will be connected to the hub with the least cost to transport it there, for example if it cost 11 dollars to transport from node 11 to hub 14 and 27 dollars to transport it to hub 2, node 11 will be connected with hub 14.
-This process is repeated until the algorithm goes through all nodes.
-To illustrate we will use the Large and small data assignment.
+        Transport to the same city is free if the city is a hub.
+To optimize the total cost an algorithm is made that will look at the city with the most flow of parcels to itself and make a hub out of this city.
+After calculating the cost, the next city with the most flow of parcels to itself will be become a hub.
+If the solution becomes cheaper by adding a hub, the city is added to a list of hubs.
+If the total cost increases, nothing happens and the next city (most parcels to itself) becomes a hub.
+If there are two cities with the same amount of parcels to itself the algorithm will choose the node with the most incoming parcels.
+Each node will be connected to the hub with the least cost to transport it there, for example if it cost €11 to transport from city 11 to hub 14 and €27 to transport it to hub 2, city 11 will be connected with hub 14.
+This process is repeated until the algorithm goes through all cities.
+To illustrate we will use the Large and Small data assignment.
 \subsubsection{Result of the large data set}
         \begin{table}[h!]
-            \begin{tabular}{|p{3cm}|p{3cm}|p{3cm}|}
+            \begin{tabular}{||c|c|c||}
                 \hline
@@
 
-The cheapest situation here is where there are 4 hubs in nodes 2,6,14 and 15.
-We connect each node with the hub where the cost to transport a parcel is the lowest, the total cost in this situation would be €156.5 thousand.
-One can easily see that it is very profitable to make a hub out of node 14, as it is the node with the most incoming parcels (610).
+The cheapest situation is when there are 4 hubs in cities 2,6,14 and 15.
+We connect each city with the hub where the cost to transport a parcel is the lowest, the total cost in this situation would be €156.5 thousand.
+One can easily see that it is very profitable to make a hub out of city 14, as it is the city with the most incoming parcels (610).
 In the following subsections we will be looking at the results of the other algorithms and combine those to see if there are similarities.
 \subsubsection{Result of the small data set}
-(The results of the small data set are still being worked on, though this will be done identically to the large data set.)
+
+        \begin{table}[h!]
+            \begin{tabular}{||c|c|c||}
+                \hline
+                nodes that are hubs & flow of parcels to itself (last node) & Total cost (€ x 10^3)\\
+                \hline
+                [2] &  39 & 63.6\\
+                \hline
+                [2,6]   & 18 & 46.5 \\
+                \hline
+                [2,6,4] & 11 & 46.3 \\
+                \hline
+                [2,6,4,3]  & 9 & 49.1  \\
+                \hline
+                [2,6,4,3,7] & 9 & 51.1 \\
+                \hline
+                [2,6,4,3,7,9] & 9 & 54.2   \\
+                \hline
+                [2,6,4,3,7,9,5] & 8 & 58.2 \\
+                \hline
+            \end{tabular}
+        \caption{Table of nodes that were used as hubs and the associated costs (Small Data)}
+        \end{table}
 
@@
 But, after adding hub 9 the total costs increase from there onward.
-So, based on these results, a solution would be to pick hubs 2, 8, 10 and 14 with a total cost of 152974.
+So, based on these results, a solution would be to pick hubs 2, 8, 10 and 14 with a total cost of €152.974.
 
@@
 \subsubsection{Result of the small data set}
-(The results of the small data set are still being worked on, though this will be done identically to the large data set.)
+For the small data set, the same algorithm was used. However, the input and programw as slightly modified to fit the small data set. The results can be seen in table \ref{DistAlgSmall1}.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{||c|c|c||}
+        \hline
+        Chosen hub & Total distribution costs (€ x 10^3)& Total costs (€ x 10^3)  \\
+        \hline
+        \hline
+        6 & 18.4 & 49.3 \\
+        \hline
+        2 & 13.6 & 46.7 \\
+        \hline
+        4 & 9.6 & 46.3 \\
+        \hline
+        5 & 7.2 & 48.4 \\
+        \hline
+        1 & 5.4 & 50.0 \\
+        \hline
+        3 & 3.7 & 53.0 \\
+        \hline
+        8 & 2.2 & 55.3 \\
+        \hline
+        7 & 1.2 & 58.0 \\
+        \hline
+        9 & 0.8 & 63.0 \\
+        \hline
+        10 & 0 & 66.3 \\
+        \hline
+    \end{tabular}
+    \caption{Results based on distribution costs for the small data set}
+    \label{DistAlgSmall1}
+\end{table}
+
+We see that in this case the optimal solution occurs when hubs 2, 4 and 6 are chosen, with a total cost of €46.330.
+Once again, the algorithm was also run based on total costs, the results of which can be seen in table \ref{DistAlgSmall2}.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{||c|c||}
+    \hline
+        Chosen hub & Total costs (€ x 10^3) \\
+        \hline
+        \hline
+        6 & 49.3 \\
+        \hline
+        2 & 46.7 \\
+        \hline
+        4 & 46.3 \\
+        \hline
+        1 & 48.0 \\
+        \hline
+        5 & 50.0 \\
+        \hline
+        7 & 52.1 \\
+        \hline
+        3 & 54.9 \\
+        \hline
+        8 & 58.0 \\
+        \hline
+        9 & 62.9 \\
+        \hline
+        10 & 66.3 \\
+        \hline
+    \end{tabular}
+    \caption{Results based on total costs for the small data set}
+    \label{DistAlgSmall2}
+\end{table}
+
+Notably; the first three cities to be added as hubs are the exact same for both cases, with the two after that being switched around.
+This might suggest that these cities are all good candidates for a hub.
+We also note that the optimal solution based on total costs is the same, namely hubs 2, 4 and 6 with a total cost of €46.330.
 
@@
     \subsubsection{Result of the large data set}
+    \label{sssec:num2}
 By combining and analysing the results of the different algorithms, a hub will be opened at nodes 14 and 6 as these nodes were also hubs in the cheapest situation of all different algorithms that were used.
-By opening a hub at node 14 the cost for collecting , transferring and distributing can be reduced, as it is the node with the most amount of incoming parcels.
+By opening a hub at node 14 the cost for collecting, transferring and distributing can be reduced, as it is the node with the most amount of incoming parcels.
 The total cost in this situation where we only have two hubs is €144.870.
@@
 \hline
+
 Nodes that are hubs & Total cost (€ x 10^3)\\
-\hline
+\hline\hline
 [6,14] &  144.8 \\
@@
 \begin{tikzpicture}[main/.style = {draw, circle,minimum size=0.8cm}]
-\node[main] (6) {6};
-\node[main] (8) [below right=1.5cm of 6] {8};
-\node[main] (14) [below left=1.5cm of 6] {14};
+\node[main] (6) [fill = blue!20] {6};
+\node[main] (8) [below right=1.5cm of 6, fill =blue!20] {8};
+\node[main] (14) [below left=1.5cm of 6, fill =blue!20] {14};
 
@@
 \subsubsection{Result of the small data set}
-(The results of the small data set are still being worked on, though this will be done identically to the large data set.)
+By combining and analysing the results of the different algorithms, a hub will be opened at nodes 2, 4 and 6.
+In the algorithms for both transfer and distribution the combination of these 3 offered the cheapest solution.
+The algorithm for collection costs does suggest that city 2 and 6 are good candidates as well, though less so for city 4.
+This results in a total cost of €46.330.
+Beyond these 3 cities, no other cities were found as viable hubs to make it even cheaper.
+
+Notably, these hubs have the largest flow to themselves.
+Making them hubs reduces the costs of collecting, transferring and distributing those parcels to €0.
+Cities 2 and 6 also have some of the most outgoing and incoming parcels, meaning a lot is saved on collection and distribution costs there.
+
+\begin{figure}[h]
+\centering
+\begin{tikzpicture}[main/.style = {draw, circle,minimum size=0.8cm}]
+\node[main] (6) [fill = blue!20] {6};
+\node[main] (2) [below right=1.5cm of 6, fill = blue!20] {2};
+\node[main] (4) [below left=1.5cm of 6, fill = blue!20] {4};
+
+\node[main] (1) [right=1cm of 2] {1};
+
+\node[main] (3) [above=of 6,xshift=-1.7cm,yshift=-1.7cm] {3};
+\node[main] (5) [above=of 6,xshift=-1.3cm,yshift=-0.9cm] {5};
+\node[main] (7) [above=of 6,xshift=-0.5cm,yshift=-0.5cm] {7};
+\node[main] (8) [above=of 6,xshift=0.5cm,yshift=-0.5cm] {8};
+\node[main] (9) [above=of 6,xshift=1.3cm,yshift=-0.9cm] {9};
+\node[main] (10) [above=of 6, xshift=1.7cm,yshift=-1.7cm] {10};
+
+\draw[-] (6) -- (4);
+\draw[-] (6) -- (2);
+\draw[-] (4) -- (2);
+
+\draw[-] (6) -- (3);
+\draw[-] (6) -- (5);
+\draw[-] (6) -- (7);
+\draw[-] (6) -- (8);
+\draw[-] (6) -- (9);
+\draw[-] (6) -- (10);
+
+\draw[-] (2) -- (1);
+\end{tikzpicture}
+\caption{A graph of the optimal solution for the small data set}
+\end{figure}
 
@@
 \subsection{ILP implementation}
+\label{ILPimp}
 For the implementation of the ILP the actual values of the inputs will not be used, rather variables to represent any valid input for the ILP.
@@
 Next, a means is required that indicates which path is valid for a package to take from a city i to j.
-A way to do this is by requiring that, for some i,j,k,l \in N, both $e_{ik} = 1$ and $e_{lj} = 1$.
+A way to do this is by requiring that, for some i,j,k,l $\in$ N, both $e_{ik} = 1$ and $e_{lj} = 1$.
 However, this would not be linear.
@@
 
-To make this variable mean the same as what is desired, the following three constraints are introduced:
-
-\begin{align*}
-    y_{ijkl} &\leq e_{ik} \\
-    y_{ijkl} &\leq e_{jl} \\
-    y_{ijkl} &\geq e_{ik} + e_{jl} - 1
-\end{align*}
-
-The validity of these constraints is shown in table \ref{tab:Helper}.
+How this helper variable is made to be the same as what is desired will be done in the constraints.
+Using these variables, the objective function and constraints are as follows:
+
+\begin{align}
+    min \, z &= \sum^N_{i=1}\sum^N_{j=1}\sum^N_{k=1}\sum^N_{l=1} w_{ij}(\chi c_{ik} + \alpha c_{kl} + \delta c_{lj})y_{ijkl} + \sum^N_{k=1} f_k e_{kk} \label{obj}\\
+    \sum^N_{k=1}e_{ik} &= 1 \quad\quad i \in N \label{C1}\\
+    e_{ik} & \leq e_{kk} \quad\quad i,k \in N \label{C2}\\
+    y_{ijkl} &\leq e_{ik} \quad\quad i,j,k,l \in N \label{C3}\\
+    y_{ijkl} &\leq e_{jl} \quad\quad i,j,k,l \in N \label{C4}\\
+    y_{ijkl} &\geq e_{ik} + e_{jl} - 1 \quad\quad i,j,k,l \in N \label{C5}\\
+    \sum^N_{k=1}e_{kk} &\geq 1 \quad\quad k \in N \label{C6}
+\end{align}
+
+The objective function (\ref{obj}) is constructed as follows: given a path from $i \rightarrow k \rightarrow l \rightarrow j$ with i,j,k,l $\in$ N and the relevant multipliers we get that sending one unit of flow over this path is $\chi c_{ik} + \alpha c_{kl} + \delta c_{lj}$.
+However, we want to send more than just one package.
+So, it is multiplied by the total flow between the sending city i and receiving city j, namely $w_{ij}$.
+To make sure that the chosen path is valid, that is city i is assigned to hub k and city j is assigned to hub l, it is also multiplied by $y_{ijkl}$.
+Finally, we sum over all the indexes i,j,k,l to get the total cost for sending packages.
+
+On top of that, there is also the costs of establishing hubs at certain cities.
+We multiply this cost for a certain city k $\in$ N by $e_{kk}$, since only hubs are connected to themselves.
+By summing over k the total costs for establishing hubs is calculated.
+
+The first constraint (\ref{C1}) is there to make sure that each city needs to be assigned to exactly one hub.
+Similarly, the second constraint (\ref{C2}) makes sure that a city i can only be assigned to a city k if city k itself is a hub.
+
+The next three constraints (\ref{C3}) (\ref{C4}) (\ref{C5}) have been made such that $y_{ijkl} = e_{ik}*e_{lj}$. The validity of this claim is shown in table \ref{tab:Helper}.
 
@@
 
-With this helper variable defined it is now possible to construct the objective function.
-First of all, the cost for a unit of flow sent along a certain path is needed.
-Given a path from $i \rightarrow k \rightarrow l \rightarrow j$ with i,j,k,l $\in$ N and the relevant multipliers we get that sending one unit of flow over this path is $\chi c_{ik} + \alpha c_{kl} + \delta c_{lj}$.
-However, we want to send more than just one package.
-So, it is multiplied by the total flow between the sending city i and receiving city j, namely $w_{ij}$.
-To make sure that the chosen path is valid, that is city i is assigned to hub k and city j is assigned to hub l, it is also multiplied by $y_{ijkl}$.
-Summing over i,j,k,l $\in$ N, the total costs for sending packages over the chosen hubs and connections is:
-$$\sum^N_{i=1}\sum^N_{j=1}\sum^N_{k=1}\sum^N_{l=1} w_{ij}(\chi c_{ik} + \alpha c_{kl} + \delta c_{lj})y_{ijkl}$$
-
-On top of that, there is also the costs of establishing hubs at certain cities.
-We multiply this cost for a certain city k $\in$ N by $e_{kk}$, since only hubs are connected to themselves.
-Summing over k, this gives the total cost of establishing the hubs as:
-$$\sum^N_{k=1} f_k e_{kk}$$
-
-Combining this with the previous part and the fact that the objective function needs to be minimised we obtain that the objective function is:
-$$min \sum^N_{i=1}\sum^N_{j=1}\sum^N_{k=1}\sum^N_{l=1} w_{ij}(\chi c_{ik} + \alpha c_{kl} + \delta c_{lj})y_{ijkl} + \sum^N_{k=1} f_k e_{kk}$$
-
-Lastly, some additional constraints are introduced.
-Firstly, since every city needs to be assigned to exactly one hub, $\sum^N_k e_{ik} = 1$ for all i $\in$ N is introduced as a constraint.
-Additionally, the constraint $e_{ik} \leq e_{kk}$ for all i,k $\in$ N is set as a constraint as well.
-For the latter, when $e_{kk} = 0$, it is a city.
-Thus, there can be no other cities assigned to it, which that constraints makes sure of.
-When $e_{kk} = 1$, it is a hub, meaning that there can be a different city assigned to it, though it is not required.
-Finally, a constraint is introduced that requires there to be at least one hub, namely $\sum^N_k e_{kk} \geq 1$.
-With these constraints the ILP definition is finalised and it can now be used to calculate an optimal solution.
+Finally, the sixth constraint (\ref{C6}) makes it so that there is at least one hub in the final answer.
 
 \subsection{ILP result}
-The ILP model was implemented into Python using PuLP, with CPLEX as solver.
-The relevant data was imported from an excel file using Pandas.
-The full program can be found in Appendix \ref{ILPPyth}.
-After running the program it returned an optimal solution of 144.369.
+The ILP model was implemented into Python using PuLP, with CPLEX as solver. The relevant data was imported from an excel file using Pandas. The full program can be found in Appendix \ref{ILPPyth}. 
+\subsubsection{Result of the large data set}
+After running the program for the large data set it returned an optimal solution of €144.369.
 Interestingly, the hubs chosen by this program were 6, 8 and 14, the exact same ones as the result from the intuitive algorithm.
-However, that resulted in a cost of 144.532, which is a difference of 163.
-
-Upon closer inspection, it was discovered that one connection differed from the intuitive algorithm.
+However, that resulted in a cost of €144.532, which is a difference of 163.
+
+Upon closer inspection, it was discovered that one connection differed from the solution of the intuitive algorithm.
 In the ILP optimal solution, a graph of which can be found in figure \ref{ILPSol}, city 10 was connected to hub 14 instead of hub 6.
@@
 \begin{tikzpicture}[main/.style = {draw, circle,minimum size=0.8cm}]
-\node[main] (6) {6};
-\node[main] (8) [below right=1.5cm of 6] {8};
-\node[main] (14) [below left=1.5cm of 6] {14};
+\node[main] (6) [fill = blue!20] {6};
+\node[main] (8) [below right=1.5cm of 6, fill = blue!20] {8};
+\node[main] (14) [below left=1.5cm of 6, fill = blue!20] {14};
 
@@
 \end{tikzpicture}
-\caption{A graph of the optimal ILP solution. The 10 coloured red indicates the result of the intuitive algoritm while the 10 coloured green indicates the result of the ILP solution.}
+\caption{\centering A graph of the optimal ILP solution for the large data set.
+The 10 coloured red indicates the result of the intuitive algorithm while the 10 coloured green indicates the result of the ILP solution.}
 \label{ILPSol}
+\end{figure}
+
+\subsubsection{Result of the small data set}
+After running the the program for the small data set it returned an optimal solution of €46.330.
+Cities 2, 4 and 6 were chosen by the program as hubs.
+Notably, these are the same cities that were chosen as hubs by the intuitive algorithm.
+In fact, the resulting optimal solution is exactly the same as well.
+The graph of this solution can be found in figure \ref{ILPSolSmall}.
+So, it appears that using the intuitive algorithm the best solution had already been found, reinforcing the accuracy of the intuitive algorithm.
+
+\begin{figure}[h]
+
+\centering
+\begin{tikzpicture}[main/.style = {draw, circle,minimum size=0.8cm}]
+\node[main] (6) [fill = blue!20] {6};
+\node[main] (2) [below right=1.5cm of 6, fill = blue!20] {2};
+\node[main] (4) [below left=1.5cm of 6, fill = blue!20] {4};
+
+\node[main] (1) [right=1cm of 2] {1};
+
+\node[main] (3) [above=of 6,xshift=-1.7cm,yshift=-1.7cm] {3};
+\node[main] (5) [above=of 6,xshift=-1.3cm,yshift=-0.9cm] {5};
+\node[main] (7) [above=of 6,xshift=-0.5cm,yshift=-0.5cm] {7};
+\node[main] (8) [above=of 6,xshift=0.5cm,yshift=-0.5cm] {8};
+\node[main] (9) [above=of 6,xshift=1.3cm,yshift=-0.9cm] {9};
+\node[main] (10) [above=of 6, xshift=1.7cm,yshift=-1.7cm] {10};
+
+\draw[-] (6) -- (4);
+\draw[-] (6) -- (2);
+\draw[-] (4) -- (2);
+
+\draw[-] (6) -- (3);
+\draw[-] (6) -- (5);
+\draw[-] (6) -- (7);
+\draw[-] (6) -- (8);
+\draw[-] (6) -- (9);
+\draw[-] (6) -- (10);
+
+\draw[-] (2) -- (1);
+\end{tikzpicture}
+\caption{A graph of the optimal ILP solution for the small data set}
+\label{ILPSolSmall}
 \end{figure}
@@
                 \begin{center}
-                    \begin{tabular}{ |p{3cm}|p{3cm}|p{4cm}|}
+                    \begin{tabular}{ |p{3cm}|p{4cm}|p{4cm}|}
                         \hline
@@
                         \hline
-                        [9,10 ]& 88.4 | [6, 10, 2, 14] & 148.8 |  [9, 10, 2, 14]\\
-                        \hline
-                        [6, 14] &  66.6 | [6, 14]  &  131.7 | [6, 14] \\
-                        \hline
-                        [6, 14, 8] & 75.6 | [2, 10, 14] & 136.3 | [6,14,8]  \\
+                        [9]& 88.4 | [2, 9,10, 14] & 151.3 |  [9, 10, 2, 14]\\
+                        \hline
+                        [2] &  111.8 | [2,3,5,9,10,14]  &  170.8 | [2,3,5,6,10,14] \\
+                        \hline
+                        [6, 14, 8] & 75.6 | [6, 14,8] & 136.3 | [6,14,8]  \\
+                        \hline
+                        [6, 14] & 66.6| [6, 14] & 131.7 | [6,14]  \\
                         \hline
@@
             \caption{Table of the amount of times each city is used as a hub in every combination}
-            \label{amount of times hubs are used in collection} 
-            \end{table}
-
-            This table shows that most likely turning cities 6 and 14 into hubs will result into a low cost. Moreover, cities 2, 10 and 8 also have a high probability of ending up in the final result.
+            \label{amount of times hubs are used in collection expanded} 
+            \end{table}
+
+            This table shows that most likely turning cities 6 and 14 into hubs will result into a low cost. Moreover, the combination of 10, 2 and 14 also gives a relatively cost efficient result.
 
@@
             \begin{table}[h!]
-            \begin{tabular}{|p{3cm}|p{3cm}|p{3cm}|}
+            \begin{tabular}{||c|c|c||}
             \hline
@@
             \hline
-            [14] &  181 & 404.3 \\
-            \hline
-            [14,2]   & 39 & 282.4 \\
-            \hline
-            [14,2,15] & 26 & 273.1 \\
-            \hline
-            [14,2,15,6]  & 18 & 228.9 \\
-            \hline
-            [14,2,15,6,13] & 18 & 222.2\\
-            \hline
-            [14,2,15,6,13,4] & 11 & 213.6   \\
-            \hline
-            [14,2,15,6,13,4,3] & 9 & 214.9   \\
-            \hline
-            [14,2,15,6,13,4,3,9] & 9 & 218.5    \\
-            \hline
-            [14,2,15,6,13,4,3,9,7] & 9 & 222.9    \\
+            \hline
+            [14] &  181 & 165.4 \\
+            \hline
+            [14,2]   & 39 & 144.5 \\
+            \hline
+            [14,2,15] & 26 & 150.7 \\
+            \hline
+            [14,2,15,6]  & 18 & 140.5 \\
+            \hline
+            [14,2,15,6,13] & 18 & 147.4\\
+            \hline
+            [14,2,15,6,13,4] & 11 & 155.4   \\
+            \hline
+            [14,2,15,6,13,4,3] & 9 & 166.4   \\
+            \hline
+            [14,2,15,6,13,4,3,9] & 9 & 178.6    \\
+            \hline
+            [14,2,15,6,13,4,3,9,7] & 9 & 189.6    \\
             \hline
@@
 
-The cheapest situation here is where we have 6 hubs in nodes 2,4,6,13,14,15, the total cost in this situation is €213.621 which is €57086 more expensive than before.
+A solution would be to pick hubs 2,6,14 and 15 with a total cost of €140565.
 An explanation for this is that the packages are not sent on one go, but in two or more times, which would ultimately lead to more costs.
-There is also a big difference in the locations of the hubs.
-First there were 4 (2,6,14 and 15) hubs and now two extra hubs are being established.
-In the following chapters the other algorithms will be implemented and combined so that in the end an optimal situation arises.
+There is no difference with the locations of the hubs in section \ref{sssec:num1}.
+In the following chapters the other algorithms will b3e implemented and combined so that in the end an optimal situation arises.
 
@@
 
-\subsection{ILP result}
-(This part of the expanded algorithm is still being worked on)
+\subsubsection{Result of the combined algorithms}
+
+As in section \ref{sssec:num2} hubs will be established in cities 6 and 14. The same algorithm used in section \ref{sssec:num2} will be implemented to find the optimal solution.
+
+\begin{table}[h]
+\begin{tabular}{||c |c||}
+\hline
+Nodes that are hubs & Total cost (€ x 10^3)\\
+\hline
+\hline
+[6,14] &  131.7 \\
+\hline
+[6,14,2]   & 133.2\\
+\hline
+[6,14,2,8] & 137.6 \\
+\hline
+[6,14,2,8,13]  & 144.4  \\
+\hline
+[6,14,2,8,13,15] & 152.5 \\
+\hline
+[6,14,2,8,13,15,12] & 161.1 \\
+\hline
+[6,14,2,8,13,15,12,5] & 170.0 \\
+\hline
+\end{tabular}
+\caption{Table of nodes that were used as hubs in the expanded algorithm}
+\end{table}
+
+The optimal solution that is found is when hubs are established in nodes 6 and 14 with a total cost of €131.668
+
+\subsection{Expanded ILP solver}
+Unfortunately, the attempt at an ILP solver for the expanded problem was unsuccessful. However, the implementation and the obtained result will be detailed below still.
+\subsection{Expanded ILP implementation}
+The expanded ILP implementation builds further upon the ILP implementation in section \ref{ILPimp}. First, we start by introducing 3 new variables:
+\begin{align}
+mc_{ij} &= 
+\end{align}
+
+$$min \, z &= \sum^N_{i=1}\sum^N_{j=1}\sum^N_{k=1}\sum^N_{l=1} w_{ij}(\chi c_{ik} + \alpha c_{kl} + \delta c_{lj})y_{ijkl}$$
+\subsection{Expanded ILP result}
 
@@
 \newpage
-\section{Bibliography}
+\section*{Bibliography}
+\addcontentsline{toc}{section}{Bibliography}
 \begin{thebibliography}{9}
@@
 \appendix
-\section{ILP definition}
+\section{Cost function Intuitive Algorithm}
+\url{https://github.com/Wessel201/Notes-Modelleren-B-20-04-2021}
+\label{CostFunction}
+\begin{verbatim}
+import pandas as pd 
+import matplotlib.pyplot as plt
+import math
+collection_factor = 1
+transfer_factor = 1
+distribution_factor = 1
+bus_capactity = 100
+
+file_name ='data_klein.xlsx'
+
+class parcel():
+    def __init__(self,file):
+        #Here the data is imported from the excel file and stored and variables to be used later
+        xls = pd.ExcelFile(file)
+        self.flow = pd.read_excel(xls, 'w')
+        self.flow = self.flow.transpose()
+        self.package_cost = pd.read_excel(xls, 'c')
+        self.package_cost = self.package_cost.transpose()
+        if file == 'Large.xlsx':
+            self.hub_cost =[13530] + list(pd.read_excel(xls, 'f')[13530])
+        else:
+            self.hub_cost =[4388] + list(pd.read_excel(xls, 'f')[4388])
+        self.cities = len(self.flow.columns)
+
+    def calculate_cost(self,hubs):
+		# hubs are given as input variable in the form of a list so eg. [1,2,5]
+        cost = 0
+        index = {} 
+        hubs = sorted(hubs)
+        flow_copy = self.flow.copy(deep=True)
+        flow_copy = flow_copy.drop(index = 'Unnamed: 0')
+        cost_copy = self.package_cost.copy(deep=True)
+        cost_copy = cost_copy.take([hub for hub in hubs])
+        # The following loop calculates the cost of collection and building of the hubs. 
+        for i in range(self.cities):
+            if i+1 in hubs:
+                #If statement checks if the hub is in the inputed if yes then the cost to built this hub is added to the total cost
+                index[i] = i+1
+                cost+= self.hub_cost[i]
+                continue
+            nearest_hub = cost_copy.loc[cost_copy[i] > 0 , i].idxmin()
+            #the nearest hub is determined by finding the hub that has the lowest collection cost for that city 
+            cost += cost_copy[i][nearest_hub]
+            flow_copy[i].sum()*math.ceil(flow_copy[i].sum()/bus_capactity)*collection_factor
+            #The amount of packages multiplied by the cost of collection of the nearest hub multiplied by the collection factor is added to the cost 
+            nearest_hub = nearest_hub 
+            index[i] = nearest_hub
+            #The city is linked to the hub 
+            flow_copy[nearest_hub-1] = flow_copy[i]+flow_copy[nearest_hub-1]
+            flow_copy = flow_copy.drop(columns = i)
+        for i in range(self.cities):
+            #in this loop the cost of transport and delivery is determined. 
+            flowss = list(flow_copy.iloc[i])
+            to_hub = index[i] - 1 
+            prices = list(cost_copy[to_hub])
+            total_packages = sum(flowss)
+            transfer_cost = sum([int(flowss[i])*int(prices[i]) for i in range(len(flowss))])
+            distribution_cost = self.package_cost[to_hub][i+1]*math.ceil(total_packages/bus_capactity)*distribution_factor*total_packages
+            cost+= distribution_cost + transfer_cost
+        return cost
+
+\end{verbatim}
+
+\section{Intuitive Algorithm Python implementation}
+Below are the 3 programs used to obtain the result of the intuitive algorithm. For each the cost function from Appendix \ref{CostFunction} is used, slightly modified where necessary.
+
+\subsection{Collection}
+\subsection{Transfer}
+\subsection{Distribution}
+As can be seen in line 1, this version uses the large data set. For the small data set, only the file name needs to be changed.
+\begin{verbatim}
+file_name = 'Large.xlsx'
+g = parcel(file_name)
+
+def next_lowest(g, hubs): #finds the next best hub to minimalise dist costs
+    low = 1e10
+    low_id = None
+
+    for c in range(1,g.cities):
+        #print('low', low, low_id)
+        if c in hubs:   #checks if c is already a hub
+            continue
+
+        else:           #calculates the cost of adding city c as hub
+            temp_hubs = sorted(hubs + [c])
+            #print(temp_hubs)
+            temp_cost = g.calculate_cost(temp_hubs)
+            #print('temp', temp_cost, c)
+            if temp_cost < low:     #compares to see if it's the lowest
+                low = temp_cost
+                low_id = c
+    return low, low_id
+
+cur_hubs = []
+res = []
+for _ in range(1, g.cities):
+    new_cost, new_hub = next_lowest(g, cur_hubs)
+    cur_hubs = sorted(cur_hubs + [new_hub])
+    res.append((new_cost, cur_hubs))
+
+print(res)
+\end{verbatim}
+\newpage
+
+\section{ILP solver}
 \label{ILPDef}

### 2021-06-17 Minutes/Minutes_Modelleren_4.8.tex

@@
 \begin{document}
-\section*{Minutes Wednesday}
-Aanwezig: Bart, Dyami, Severijn
-Voorzitter: Dyami
+\section*{Minutes Tuesday}
+Aanwezig: Bart, Dyami, Severijn \\
+Voorzitter: Dyami \\
 Secretaris: Bart
@@
 \end{itemize}
+
+\section*{Minutes Thursday}
+Aanwezig: Iedereen \\
+Voorzitter: Dyami \\ 
+Secretaris: Bart \\
+
+\subsection*{Discussed topics}
+\begin{itemize}
+    \item Discuss the received feedback on the draft report.
+    \item Discuss the pitch made by Severijn
+\end{itemize}
+
+\subsection*{Decisions}
+\begin{itemize}
+    \item Report needs to be done by Tuesday so that we can discuss it by Thursday. 
+\end{itemize}
+
+\subsection*{Planning}
+\begin{itemize}
+    \item Incorporate feedback into report by Tuesday (Everyone)
+    \item Add small data set results to report (Everyone)
+    \item Finalize Pitch (Severijn)
+    \item Expanded problem ILP implementation by tomorrow (Bart)
+\end{itemize}
 \end{document}

### 2021-06-17 Reports/Evaluation_Report.tex
Updates from Overleaf

\documentclass{article}
\usepackage[utf8]{inputenc}

\title{Evaluation Report Modelleren-B}
\author{Bart Stolk \and Dyami Silvério \and Wessel van Sommeren \and Severijn Wiechers}
\date{03 June 2021}

\begin{document}

\maketitle
\newpage
\tableofcontents
\newpage

\section{Introduction}

In this project we were asked to optimize the cost of a parcel delivery network for Post NL using a data set provided. The team consisted of Wessel van Sommeren, Dyami Silvério, Bart Stolk and Severijn Wiechers. We first focused on making an intuitive algorithm to solve the problem. WE did this by splitting up the problem into sub problems and optimizing those. Once we found a good solution for the problem we moved to a Integer Linear Programming implementation of the problem. The code was written in a way that everyone helped to design the algorithm and then one person would write the code. If there were any bugs or complications the rest of the team would help to solve them. The last part of the project consisted of expanding the problem. Once we agreed on how to expand the problem we would start the cycle again, with first looking at a intuitive approach followed by and ILP implementation.

2. Planning: What was the initial planning? How was it adapted? Did you follow
the planning? What did you do differently? Was the planning useful? Looking
back, what could you have planned better?
\section{Planning}

The initial plan was to meet with each other every Tuesday and Thursday. On Tuesdays we would discuss and work on ideas for the project and on Thursday work was divided between the four of us. This process went well and everyone kept to the agreements. One some days one of us wouldn't be able to show up to the agreed time due to for instance a doctor's appointment, however this was reported a day before to the whole group. The planning generally went well, as everything was completed in time.

\section{Project management}

    The planning and managing of the group and project was done at the end of meetings. Plans were made for what had to be done by next time or some other deadline. On a number of ocassions one or more of the group members were absent, but they either already know what to do for next time or the planning was communicated to them through group chats/the minutes. Some deadlines had to be moved around over the course of the project due to certain parts not always being done in time. However, this never was a cause for great concern, as eventually everything was completed in time.

4. File management and communication: How did you store and manage the files?
How did you communicate? Did problems occur? What could have been done
better?

\section{File management}

    Over the course of the writing of the report we used two main sources for storing data. Firstly, all data sets, Python scripts and report were stored in github. Since most of the group was not familiar with this program, the first day was used to explore git. In hindsight, this was quite useful since github gives many opportunities when working in a group. Furthermore, all the reports in git were linked to overleaf, the second source for data storage. Using \LaTeX as a text editor we were able to give a structured and clean end report.

    Every piece of data was divided into a couple categories. Hereby we differentiated between minutes, reports and Python files. The latter divided into files from the ILP and the intuitive algorithm. By sorting all files into different subgroups we were able to create organized maps such that everything we have written down could easily be retrieved.

\section{Quality}

    There was no rigid measures by which we tested quality. However, most products made during the course of the project were still tested relatively thoroughly. The cost function was checked for quality multiple times, discovering a number of issues. For example, at one point the function gave different results for the same hubs in a different order. The ILP results were always checked to see if they made sense and compared to the results of the cost function to see if they were better at all. Lastly, reports were checked by varying team members at varying stage of writing them, taking out spelling and style errors where necessary.

\section{Evaluations}
6. Evaluations of each others work: Every group member again fills out the table
with five criteria for each of the group members. Looking back, were these five
criteria well chosen?
    \begin{itemize}
       \item Punctuality
       \item Motivation / Participation
        \item Adaptability
        \item Quality of the work 
        \item Communication
    \end{itemize}

\begin{table}[]
    \centering
    \begin{tabular}{|c|c|c|c|c|}
        \hline
        - & Bart & Dyami & Wessel & Severijn \\
        \hline
        Punctuality & & & & \\
        \hline
        Motivation
    \end{tabular}
\end{table}

\section{Personal experiences}
7. Personal experiences: Every group member writes approximately half a page on
his/her experiences in the project: What are your experiences in working in this
group? What have you learned? What did you like, and what not? Was it like
you expected, or not? What could you have done better? What do you want to
do differently next time?
\subsection{Bart Stolk}

\subsection{Dyami Silvério}

\subsection{Wessel van Sommeren}

\subsection{Severijn Wiechers}

8. Conclusions (for the whole group).
\section{Conclusions}

\end{document}

### 2021-06-17 Reports/draft_report.tex
Updates from Overleaf

@@
                     \end{tabular}
+                    \caption{Table of the amount of times each city is used as a hub in every combination small data set}
                 \end{center}
-            \caption{Table of the amount of times each city is used as a hub in every combination small data set}
             \label{amount of times hubs are used in collection small} 
@@
 To illustrate we will use the Large and Small data assignment.
+
 \subsubsection{Result of the large data set}
@@
         \end{table}
+
+The most profitable solution is to pick hubs in nodes 2,4 and 6.
 
@@
 \begin{table}[h]
+\begin{center}
 \begin{tabular}{||c |c||}
@@
 \end{tabular}
+\end{center}
 \caption{Table of nodes that were used as hubs in the expanded algorithm}
@@
 \subsection{Expanded ILP implementation}
-The expanded ILP implementation builds further upon the ILP implementation in section \ref{ILPimp}. First, we start by introducing 3 new variables:
+The expanded ILP implementation builds further upon the ILP implementation in section \ref{ILPimp}.
+First, we want the multipliers in the objective function to be variable instead of constant.
+Since the total flow over an edge in a specific direction can be split up into three categories (city to hub, hub to hub or hub to city), three new variables are introduced:
 \begin{align}
-mc_{ij} &= 
+    mc_
 \end{align}
@@
 bus_capactity = 100
+#Bus capacity is only relevant for the expanded algorithm
 

### 2021-06-17 Intuitive Algoritm/Modelleren-B Distribution Bart Stolk.py
Create Modelleren-B Distribution Bart Stolk.py

@@ -1,0 +1,87 @@
+import pandas as pd 
+collection_factor = 3
+transfer_factor = 1
+distribution_factor = 2 
+
+file_name ='Small.xlsx'
+class parcel():
+	def __init__(self,file):
+		xls = pd.ExcelFile(file)
+		self.flow = pd.read_excel(xls, 'w')
+		self.flow = self.flow.transpose()
+		self.package_cost = pd.read_excel(xls, 'c')
+		self.package_cost = self.package_cost.transpose()
+		if file == 'Large.xlsx':
+			self.hub_cost =[13530] + list(pd.read_excel(xls, 'f')[13530])
+		else:
+			self.hub_cost =[4388] + list(pd.read_excel(xls, 'f')[4388])
+		self.cities = len(self.flow.columns)
+
+	def calculate_cost(self,hubs):
+		cost = 0
+		index = {}
+		hubs = sorted(hubs)
+		flow_copy = self.flow.copy(deep=True)
+		flow_copy = flow_copy.drop(index = 'Unnamed: 0')
+		cost_copy = self.package_cost.copy(deep=True)
+		cost_copy = cost_copy.take([hub for hub in hubs])
+		for i in range(self.cities):
+			if i+1 in hubs:
+				index[i] = i+1
+				cost+= self.hub_cost[i]
+				continue
+			nearest_hub = cost_copy.loc[cost_copy[i] > 0 , i].idxmin()
+			cost += cost_copy[i][nearest_hub] * flow_copy[i].sum()*collection_factor
+			nearest_hub = nearest_hub 
+			index[i] = nearest_hub
+			flow_copy[nearest_hub-1] = flow_copy[i]+flow_copy[nearest_hub-1]
+			flow_copy = flow_copy.drop(columns = i)
+		for i in range(self.cities):
+			# print('index = ' + str(i))
+			flowss = list(flow_copy.iloc[i])
+			to_hub = index[i] - 1 
+			prices = list(cost_copy[to_hub])
+			total_packages = sum(flowss)
+			transfer_cost = sum([int(flowss[i])*int(prices[i]) for i in range(len(flowss))])
+			distribution_cost = self.package_cost[to_hub][i+1]*total_packages*distribution_factor
+			# print(distribution_cost)13530
+
+			# print(transfer_cost)
+			cost+= distribution_cost + transfer_cost
+		return cost
+
+
+		print(flow_copy)
+
+
+file_name = 'Small.xlsx'
[…]
